---
title: "Diversity-Turnover Analyses"
author: "Evan Batzer"
output:
  html_document:
    df_print: paged
---

# Part 2: Environmental Covariates


```{r, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8.5, fig.height = 5)

# Loading necessary packages
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lme4) # Linear mixed effects models
library(vegan) # General ecology functions
library(data.table) # Fast reading of data tables
library(testthat) # Unit testing
library(dtplyr) # Data.table dplyr functions
library(JostDiv) # Exponentiated diversity indices (Hill Numbers)
library(gridExtra) # Multiple GGplots
library(corrplot) # Correlation plots
```
    
### Reading in Datasets

```{r}
# Plot descriptions
plot.descriptions <- read.csv("../data/comb-by-plot-clim-soil-diversity-21-Feb-2019.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# Soil data
soil.chars <- read.csv("../data/soil-nutrients-21-February-2019.csv",
                       stringsAsFactors = FALSE,
                       na.strings = c('NA','NULL'))
# Annual climate data
climate <- read.csv(file = "../Data/Weather_annual_20190624.csv.csv")

# Loading in dataset (Updated to Feb 22st 2019)
cover <- fread('../Data/full-cover-22-February-2019.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

# Stoichiometry table
stoichrat <- read.csv("../Data/stoichtable.csv",
         stringsAsFactors = FALSE)
```

# Defining variables of interest:

Site-level:

* % Annual Taxa
* elevation
* managed
* burned
* grazed
* anthropoghenic
* latitude

Spatial resources:

* Mean C
* Mean N
* Mean P
* Mean K
* CV C
* CV N
* CV P
* CV K

Temporal resources:

* MAT
* MAP
* PPT - PET
* CV MAT
* CV MAP
* CV PPT - PET

```{r}
# Defining % Annual Taxa
ann_cov <- cover %>% 
  filter(year_trt == 0 & live == 1) %>%
  group_by(site_code) %>%
  summarise(total_cover = sum(max_cover),
            ann_cover = sum(max_cover[local_lifespan == "ANNUAL"])) %>%
  mutate(ann_frac = ann_cover / total_cover) %>%
  replace_na(list(ann_frac = 0))

# General site characteristics
site_summary <- plot.descriptions %>% 
  filter(year_trt == 0) %>%
  select(site_code, elevation, managed, burned, grazed, anthropogenic, latitude, MAP_v2, MAT_v2) %>%
  distinct()
            
# Joining together
site_summary = site_summary %>% inner_join(ann_cov %>% select(site_code, ann_frac))

# Plotting correlations
sitecor = cor(site_summary %>% select(-site_code) %>% na.omit())
corrplot(sitecor, method="circle")

# VIF Values
diag(solve(sitecor))
```

Notes:

Soil C and N concentrations seem to be highly correlated with one another (VIF ~ 5)

```{r}
# Subsetting species matrix by these site codes
soil.chars = soil.chars %>% 
  filter(year_trt == 0) %>%
  mutate(block = as.factor(block),
         plot = as.factor(plot))

View(soil.chars %>% filter(site_code == "mcla.us"))

# Selecting parameters of interest
soil.stdized = soil.chars %>% 
  select(site_code, pct_N, pct_C, ppm_P, ppm_K, pH) %>%
  na.omit() %>%
  group_by(site_code) %>%
  summarise(mean_N = mean(pct_N),
            mean_C = mean(pct_C),
            mean_P = mean(ppm_P),
            mean_K = mean(ppm_K),
            mean_pH = mean(pH),
            sd_N = sd(pct_N),
            sd_C = sd(pct_C),
            sd_P = sd(ppm_P),
            sd_K = sd(ppm_K),
            var_pH = sd(pH)) %>%
  mutate(cv_N = sd_N / mean_N,
         cv_C = sd_C / mean_C,
         cv_P = sd_P / mean_P,
         cv_K = sd_K / mean_K) %>%
  mutate_if(is.numeric, log) %>%
  filter_all(any_vars(is.finite(.))) %>%
  ungroup() %>%
  select(-contains("sd")) %>%
  rename("sd_pH" = "var_pH")
 
soil.stdized %>%
  gather(key = "var", value = "val", -site_code) %>%
  ggplot(aes(x = val)) +
  geom_histogram() +
  facet_wrap(~var, scales = "free")

soil.stdized <- soil.stdized[!is.infinite(rowSums(soil.stdized %>% select(-site_code))),]
soilcor = cor(soil.stdized %>% na.omit() %>% select(cv_N, cv_P, cv_C, cv_K,
                                      mean_N, mean_P, mean_C, mean_K, mean_pH, sd_pH))
corrplot(soilcor, method="circle")
diag(solve(soilcor))
```


```{r}
climate_summary = climate %>%
  filter(year < 2019 & 
           ppt > 0 &
           pet > 0) %>%
  group_by(site_code) %>%
  summarize(mean_wtrdef = mean(pet - ppt),
            sd_wtrdef = sd(pet - ppt))

climcor = cor(climate_summary %>% select(mean_wtrdef, sd_wtrdef))
corrplot(climcor, method="circle")
diag(solve(climcor))
```

```{r}
stoichsummary = stoichrat %>% group_by(site_code) %>%
  summarise(mean_supply = mean(a),
            sd_supply = sd(a))
```

# Joining everything together

```{r}
full_table = full_join(site_summary, soil.stdized) %>% 
  full_join(stoichsummary) %>% 
  full_join(climate_summary)

write.csv(x = full_table ,
          "../Data/divturnover_explanatory.csv")
```


######## Old Code ####################################

# Joining everything together

```{r}
full_table = inner_join(site_summary, soil.stdized %>% group_by(site_code) %>% summarise_if(is.numeric, mean)) %>%
  left_join(climate_summary) %>%
  select(-contains("sd"))

fullcor = cor(full_table %>% select(-site_code, -mean_PPT, -mean_PET, -cv_PPT, -cv_PET, -mean_C, -cv_C) %>% na.omit())
corrplot(fullcor, type = "upper")
sort(diag(solve(fullcor)))
```

```{r}
# Function to calculate network-wide site z-scores
calc_zscores <- function(x){(x - mean(na.omit(x))) / sd(na.omit(x))}

# Z-scoring variation
soil.zscores = soil.stdized %>% 
  mutate_if(is.numeric, calc_zscores)

full_table = full_table %>%
  mutate(managed = as.factor(managed),
         burned = as.factor(burned),
         grazed = as.factor(grazed),
         anthropogenic = as.factor(anthropogenic))

spatial_model = spatdiv %>% group_by(site_code, block) %>%
  #summarise(alpha = mean(alpha),
  #          beta = mean(beta)) %>%
  left_join(full_table  %>% 
              select(-mean_PPT, -mean_PET, -cv_PPT, -cv_PET, -mean_C, -cv_C) %>%
  mutate_if(is.numeric, calc_zscores))

temporal_model = tempdiv %>% group_by(site_code, block) %>%
  #summarise(alpha = mean(alpha),
  #          beta = mean(beta)) %>%
  left_join(full_table  %>% 
              select(-mean_PPT, -mean_PET, -cv_PPT, -cv_PET, -mean_C, -cv_C) %>%
  mutate_if(is.numeric, calc_zscores))

unique(na.omit(spatial_model)$site_code) # 67 sites with spatial data
unique(na.omit(temporal_model)$site_code) # 37 sites with temporal data (likely to increase after this summer's data is added? Or can reduce duration)

unique(na.omit(temporal_model)$site_code) %in% unique(na.omit(spatial_model)$site_code)
```


```{r}
library(lmerTest)

mod_alpha = lmer(scale(alpha_q0) ~ ann_frac + mean_N + mean_P + mean_K + MAT_v2 + MAP_v2 + (1|site_code/block), data = spatial_model)
mod_beta_spatial = lmer(scale(beta_q0) ~ ann_frac + mean_N + mean_P + mean_K + MAT_v2 + MAP_v2 + (1|site_code/block), data = spatial_model)
mod_beta_temporal = lmer(scale(beta_q0) ~ ann_frac + mean_N + mean_P + mean_K + MAT_v2 + MAP_v2 + (1|site_code), data = temporal_model)

anova(mod_alpha)

library(sjPlot)

# draw a dot-and-whisker plot
p1 = plot_model(mod_alpha) +
  ggtitle("Plot Richness")
p2 = plot_model(mod_beta_spatial) +
  ggtitle("Spatial Turnover")
p3 = plot_model(mod_beta_temporal) +
  ggtitle("Temporal Turnover")

grid.arrange(p1,p2,p3, nrow = 1)
str(p2)

bind_rows(p1$data,
          p2$data,
          p3$data) %>%
  mutate(lev = c(rep("Plot Richness", 6), rep("Spatial Turnover", 6), rep("Temporal Turnover", 6)),
         alpha_val = as.numeric(if_else(p.value < 0.05, 1, .3))) %>%
  ggplot(aes(x = estimate,
             y = term,
             color = estimate > 0,
             fill = estimate > 0,
             alpha = I(alpha_val))) + 
  geom_vline(xintercept = 0) +
  geom_segment(aes(x = conf.low,
                 xend = conf.high,
                 y = term,
                 yend = term),
               size = 1) +
  geom_point(size = 5, pch = 21, color = "black") +
  facet_wrap(~lev) +
  theme_bw() +
  xlab("Standardized Coefficient Estimate") +
  ylab("Coefficient Name") +
  guides(color = FALSE,
         fill = FALSE) +
  ggtitle("Model Coefficient Effects")
 
ggsave("../Figures/model_coefficients.pdf", height = 4, width = 7)
```

```{r}
meandisp = function(x, cols){return(mean(dist(data.frame(x)[,cols])))}

blockdisp = soil.chars %>% 
  select(site_code, block, pct_N, pct_C, ppm_P, ppm_K) %>%
  na.omit() %>%
  group_by(site_code, block) %>%
  mutate_if(is.numeric, scale)%>%
  do(data.frame(meandisp(., c(2:5))))

blocksupp = soil.chars %>% 
  select(site_code, block, pct_N, pct_C, ppm_P, ppm_K) %>%
  na.omit() %>%
  mutate_if(is.numeric, scale) %>%
  mutate(tot = sqrt(pct_N^2 + pct_C^2 + ppm_P^2 + ppm_K^2)) %>%
  group_by(site_code, block) %>%
  summarise(totsupply = mean(tot))

left_join(blockdisp, blocksupp)

```



# Relationships with productivity

# How to define soil resource supply?

* Supply and heterogeneity of N, P, and K?

* Resource ratio hypothesis? Variation in resource ratio?

    * Probably less appropriate, given that variance in resource supply =/= heterogeneity in resources -- doesn't account for environments that have skewed resources in different ways.
    
    
Steps:

Create aggregate data products of:

* Plot - by - plot Z-scored concentrations of limiting nutrients (C, N, P, K, MAP, MAT, PPT - PET)

* Plot pairwise interactions between both absolute concentrations of these nutrients, as well as their variance, in relation to alpha, beta, and gamma diversity, as well as productivity

# Standardizing soil data

```{r}
# Subsetting species matrix by these site codes
soil.chars = soil.chars %>% 
  filter(year_trt == 0) %>%
  mutate(block = as.factor(block),
         plot = as.factor(plot))

# Selecting parameters of interest
soil.stdized = soil.chars %>% 
  select(site_code, block, plot, pct_N, pct_C, ppm_P, ppm_K) %>%
  na.omit() %>%
  mutate_if(is.numeric, log) 

# Function to calculate network-wide site z-scores
calc_zscores <- function(x){(x - mean(na.omit(x))) / sd(na.omit(x))}

# Z-scoring variation
soil.zscores = soil.stdized %>% 
  mutate_if(is.numeric, calc_zscores)
```

# Plotting plot resource availability

```{r}
p1 = soil.zscores %>% ggplot(aes(x = pct_N)) +
  geom_histogram(color = "black", fill = "lightblue") +
  xlim(0, 1)+
  ggtitle("Plot Soil %N")

p2 =  soil.zscores %>% ggplot(aes(x = ppm_P)) +
  geom_histogram(color = "black", fill = "lightgreen") +
  ggtitle("Plot Soil ppm P")

p3 = soil.zscores %>% ggplot(aes(x = ppm_K)) +
  geom_histogram(color = "black", fill = "pink") +
  ggtitle("Plot Soil ppm K")

p4 = soil.zscores %>% ggplot(aes(x = pct_C)) +
  geom_histogram(color = "black", fill = "orange") +
  ggtitle("Plot Soil pct C")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)
```

# How to define climatic variation?

* Total MAP / MAT (absolute quantity)

* Some metric such as ETo / SPEI

# Standardizing precipitation data

```{r}
climate = climate %>%
  filter(year < 2019 & 
           ppt > 0 &
           pet > 0) %>%
  mutate(pet.z = calc_zscores(log(pet)),
         ppt.z = calc_zscores(log(ppt)))
```

```{r}
p1 <- climate %>% 
  ggplot(aes(x = ppt.z)) +
  geom_histogram(color = "black", fill = "lightblue") +
  ggtitle("Site Log(PPT)")
  
p2 <- climate %>% 
  ggplot(aes(x = pet.z)) +
  geom_histogram(color = "black", fill = "lightgreen") +
  ggtitle("Site Log(PET)")
  
p3 <- climate %>% 
  ggplot(aes(x = ppt - pet)) +
  geom_histogram(color = "black", fill = "pink") +
  ggtitle("Site PPT - PET")
  
p4 <- climate %>% 
  ggplot(aes(x = ppt.z - pet.z)) +
  geom_histogram(color = "black", fill = "orange") +
  ggtitle("Site Log(PPT) - Log(PET)")
  
gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)

```

# What best predicts alpha diversity?

```{r}
spatial_table = left_join(spatdiv, soil.zscores %>% mutate(block = as.numeric(block),
                                            plot = as.numeric(plot)),
          by = c("site_code", "block", "plot")) %>%
  left_join(plot.descriptions %>%
              group_by(site_code) %>%
              summarise_at(c("RAIN_PET", "MAT_v2", "TEMP_VAR_v2", "MAP_v2"), mean))

depvar_plot = function(obsvar, predvar, mydf){
  plot = ggplot(aes_string(x = obsvar,
           y = predvar,
           color = "site_code"),
           data = mydf) +
  geom_point(alpha = .2) +
  geom_line(stat = "smooth", method = "lm") +
  geom_line(stat = "smooth", method = "lm", aes(color = FALSE), color = "black", size = 2) +
  guides(color = FALSE)
  
  return(plot)
}

p1 = depvar_plot("pct_N", "alpha", spatial_table)
p2 = depvar_plot("pct_C", "alpha", spatial_table)
p3 = depvar_plot("ppm_P", "alpha", spatial_table)
p4 = depvar_plot("ppm_K", "alpha", spatial_table)
p5 = depvar_plot("TEMP_VAR_v2", "alpha", spatial_table)
p6 = depvar_plot("RAIN_PET", "alpha", spatial_table)
p7 = depvar_plot("MAT_v2", "alpha", spatial_table)
p8 = depvar_plot("MAP_v2", "alpha", spatial_table)

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8,
                        nrow = 2)
```

```{r}
st_blocks = spatial_table %>% 
  group_by(site_code, block) %>%
  summarise_if(is.numeric, mean)
  
p1 = depvar_plot("pct_N", "beta", st_blocks) + ylim(1, 5)
p2 = depvar_plot("pct_C", "beta", st_blocks) + ylim(1, 5)
p3 = depvar_plot("ppm_P", "beta", st_blocks) + ylim(1, 5)
p4 = depvar_plot("ppm_K", "beta", st_blocks) + ylim(1, 5)
p5 = depvar_plot("TEMP_VAR_v2", "beta", st_blocks) + ylim(1, 5)
p6 = depvar_plot("RAIN_PET", "beta", st_blocks) + ylim(1, 5)
p7 = depvar_plot("MAT_v2", "beta", st_blocks) + ylim(1, 5)
p8 = depvar_plot("MAP_v2", "beta", st_blocks) + ylim(1, 5)

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8,
                        nrow = 2)
```

```{r}
st_blocks = spatial_table %>% 
  group_by(site_code, block) %>%
  summarise(beta = mean(beta),
            N_var = var(pct_N),
            C_var = var(pct_C),
            P_var = var(ppm_P),
            K_var = var(ppm_K))
  
p1 = depvar_plot("N_var", "beta", st_blocks) + ylim(1, 5)
p2 = depvar_plot("C_var", "beta", st_blocks) + ylim(1, 5)
p3 = depvar_plot("P_var", "beta", st_blocks) + ylim(1, 5)
p4 = depvar_plot("K_var", "beta", st_blocks) + ylim(1, 5)

gridExtra::grid.arrange(p1, p2, p3, p4,
                        nrow = 2)
```

```{r}
library(lmerTest)
mod = glmer(alpha ~ scale(pct_N) + scale(pct_C) + scale(MAT_v2) + scale(MAP_v2) + scale(RAIN_PET) +  (1|site_code/block),
            data = spatial_table,
            family = "poisson")


mod = glmer(beta ~ N_var + C_var + P_var + K_var +  (1|site_code/block),
            data = st_blocks)


summary(mod)
anova(mod,type = 3)
plot(resid(mod) ~ predict(mod))
plot(mod)
```



```{r}
biomass_bysite = biomass_data %>%
  filter(year_trt == 0) %>%
  group_by(site_code, block) %>%
  mutate(total_mass = as.numeric(total_mass)) %>%
  summarise(meanprod = mean(na.omit(total_mass)),
            cvprod = sd(na.omit(total_mass)) / mean(na.omit(total_mass)),
            type = unique(habitat)) %>%
  filter(cvprod != 0 &
           meanprod != 0) 

spat_bmass = left_join(spatial_table, biomass_bysite)

p1 = depvar_plot("meanprod", "alpha", spat_bmass)
p2 = depvar_plot("log(cvprod)", "alpha", spat_bmass)
p3 = depvar_plot("meanprod", "beta", spat_bmass) + ylim(1,5)
p4 = depvar_plot("log(cvprod)", "beta", spat_bmass) + ylim(1,5)
gridExtra::grid.arrange(p1, p2, p3, p4,
                        nrow = 2)
```

