---
title: "Diversity-Turnover Analyses"
author: "Evan Batzer"
output:
  html_document:
    df_print: paged
---

### Steps to Analysis:

Subset into 2 separate datasets

1) Spatial datasets, consisting of all observational data with soil resource measurements (pretreatment only)

2) Temporal datasets, consisting of control plots with *5 years of data*, soil resource measurements.

    2a) Find those sites that contain climatic variables from Siddharth
    
3) Calculate alpha, beta, and gamma diversity between each site

    3a) Do so in a flexible way that can allow for different sorts of metrics, not just richness, but also ENS.
    
### Loading packages

```{r, message = FALSE, echo = FALSE}
# Loading necessary packages
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lme4) # Linear mixed effects models
library(vegan) # General ecology functions
library(data.table) # Fast reading of data tables
library(testthat) # Unit testing
library(dtplyr) # Data.table dplyr functions
library(JostDiv) # Exponentiated diversity indices (Hill Numbers)
library(gridExtra)
```
    
### Reading in Datasets

```{r}
# Biomass data
biomass_data <- fread("../Data/comb-by-plot-22-February-2019.csv",
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

# Loading in dataset (Updated to Feb 22st 2019)
cover <- fread('../Data/full-cover-22-February-2019.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

# Plot descriptions
plot.descriptions <-  read.csv("../Data/comb-by-plot-clim-soil-diversity-22-Feb-2019.csv",
                              stringsAsFactors = FALSE)

# Soil characteristics
soil.chars <- read.csv("../Data/soil-nutrients-21-February-2019.csv",
                              stringsAsFactors = FALSE)

# Site covariates
site.covars <- read.csv("../Data/site_covars.csv",
                        stringsAsFactors = FALSE)

```

### Filtering

```{r}
sites_wsoil <- plot.descriptions %>% 
  filter(!is.na(pct_C) & !is.na(pct_N)) %>%
  distinct(site_code)

sitevec = c(sites_wsoil$site_code)
```

# Spatial diversity
  
```{r}
com_spatial <- cover %>% 
  filter(year_trt == 0 & site_code %in% sitevec) %>% 
  filter(live == 1)

cover.wide <- data.frame(com_spatial) %>%
  select(site_code, block, plot, Taxon, max_cover) %>%
  dcast(site_code+block+plot ~ Taxon,
                    value.var='max_cover', 
                    fun.aggregate = sum,drop=T,fill=0)

calc_div = function(x, q){
  
  sample_p = vegan::decostand(x, method = "total")
  alpha = apply(sample_p, MARGIN = 1,
                d_calc, 
                q, 
                sample.weight = 1)
  
  comm_p = vegan::decostand(colSums(x), method = "total")
  gamma = d_calc(comm_p, q, sample.weight = 1)
  
  output = data.frame(alpha = alpha,
                      gamma = rep(gamma, length(alpha)),
                      beta = alpha / gamma)
  
  return(output)
  
}

spat_div = cover.wide %>% group_by(site_code, block) %>%
  do(calc_div(.[,-c(1:3)], q = 0))

write.csv("../Data/diversity_spatial_full.csv",
          x = spat_div)

spat_div = spat_div %>%
  summarise(alpha = mean(alpha),
            beta = mean(beta),
            gamma = mean(gamma))
```

### Error checking & figures

```{r}
summary(spat_div)
```

```{r, fig.height=6, fig.width=10}
p1 = spat_div %>% 
  ggplot(aes(x = alpha)) +
  geom_histogram() +
  ggtitle("Spatial alpha (plot) diversity")

p2 = spat_div %>% 
  ggplot(aes(x = beta)) +
  geom_histogram() +
  ggtitle("Spatial beta (plot \ block) diversity")

p3 = spat_div %>% 
  ggplot(aes(x = gamma)) +
  geom_histogram() +
  ggtitle("Spatial gamma (block) diversity")

grid.arrange(p1, p2, p3, nrow = 1)
```

```{r, fig.height=6, fig.width=10}
p1 = spat_div %>%
  ggplot(aes(x = alpha, y = gamma, color = site_code)) +
  geom_point(alpha = .2) +
  guides(color = FALSE) +
  stat_smooth(aes(color = FALSE), method = "lm")

p2 = spat_div %>%
  ggplot(aes(x = alpha, y = beta, color = site_code)) +
  geom_point(alpha = .2) +
  guides(color = FALSE) +
  stat_smooth(aes(color = FALSE), method = "lm")

p3 = spat_div %>%
  ggplot(aes(x = gamma, y = beta, color = site_code)) +
  geom_point(alpha = .2) +
  guides(color = FALSE) +
  stat_smooth(aes(color = FALSE), method = "lm")

grid.arrange(p1, p2, p3, nrow = 1)

```

## Temporal Diversity

To capture temporal diversity, I examine the amount of turnover present within the first 5 years of data 

```{r}
sites_temp = cover %>% select(site_code, year_trt) %>%
  group_by(site_code) %>%
  summarise(maxyr = max(year_trt),
            minyr = min(year_trt)) %>%
  filter(maxyr >= 4 & minyr == 0)

com_temporal <- cover %>% 
  filter(year_trt < 5 & trt == "Control" & site_code %in% c(sites_temp$site_code)) %>% 
  filter(live == 1)

cover.wide <- data.frame(com_temporal) %>%
  select(site_code, block, plot, year_trt, Taxon, max_cover) %>%
  dcast(site_code+block+plot+year_trt ~ Taxon,
                    value.var='max_cover', 
                    fun.aggregate = sum,drop=T,fill=0)

enoughcoverage = cover.wide %>% group_by(site_code, block, plot) %>%
  summarise(count = length(unique(year_trt))) %>%
  filter(count == 5)

calc_div = function(x, q){
  
  sample_p = vegan::decostand(x, method = "total")
  alpha = apply(sample_p, MARGIN = 1,
                d_calc, 
                q, 
                sample.weight = 1)
  
  comm_p = vegan::decostand(colSums(x), method = "total")
  gamma = d_calc(comm_p, q, sample.weight = 1)
  
  output = data.frame(alpha = alpha,
                      gamma = rep(gamma, length(alpha)),
                      beta = alpha / gamma)
  
  return(output)
  
}

temp_div = right_join(cover.wide, enoughcoverage) %>% 
  select(-count) %>% 
  group_by(site_code, block, plot) %>%
  do(calc_div(.[,-c(1:4)], q = 0)) 

write.csv("../Data/diversity_temporal_full.csv",
          x = temp_div)

temp_div = temp_div %>%
  summarise(alpha = mean(alpha),
            beta = mean(beta),
            gamma = mean(gamma))
```

```{r, fig.height=6, fig.width=10}
p1 = temp_div %>% 
  ggplot(aes(x = alpha)) +
  geom_histogram() +
  ggtitle("Temporal alpha (plot) diversity")

p2 = temp_div %>% 
  ggplot(aes(x = beta)) +
  geom_histogram() +
  ggtitle("Temporal beta (plot \ block) diversity")

p3 = temp_div %>% 
  ggplot(aes(x = gamma)) +
  geom_histogram() +
  ggtitle("Temporal gamma (block) diversity")

grid.arrange(p1, p2, p3, nrow = 1)
```

# Are temporal and spatial beta diversity correlated to one another?

```{r}
df1 = temp_div %>% rename("temp_alpha" = alpha,
                    "temp_beta" = beta,
                    "temp_gamma" = gamma)

df2 = spat_div %>% rename("spat_alpha" = alpha,
                    "spat_beta" = beta,
                    "spat_gamma" = gamma)

full_join(df1, df2) %>%
  ggplot(aes(x = temp_beta,
             y = spat_beta, 
             color = site_code)) +
  geom_point(alpha = .4) +
  guides(color = FALSE) +
  stat_smooth(aes(color = FALSE), method = "lm") +
  geom_line(stat = "smooth",
            method = "lm", se = FALSE, alpha = .2) +
  ylim(.2, .8)

```

