---
title: "R Notebook"
output: html_notebook
---

A suggested way to tease apart what makes certain species more likely to respond 
positively to different nutrient enrichment treatments is through some variety 
of decision tree.

When we have a set of hierarchical predictors that range from site-level predictors
(precip, temp, species richness) to more individual-level predictors (functional 
type, relative rank abundance), decision trees can offer a reasonably intuitive 
way to model and interpet interactions between them.

# What features of a community best predict response?

- Pretrt abundance
    - Assign a zero if missing
- Functional type
- Exotic / native status
- Any sort of other functional characteristics worth adding?

# Check if previous code has any solutions

```{r}
library(tidyverse);library(data.table);library(dtplyr)

specscores_full <- read.csv("../Data/specscores.csv")

 taxonomy <- fread('C:/Users/ebatz/Dropbox/NutNet Data/Nutnet-site-taxonomy.csv',
                stringsAsFactors = FALSE,
                na.strings = c('NA','NULL'))

cover <- fread('C:/Users/ebatz/Dropbox/NutNet Data/full-cover-22-February-2019.csv')

taxonomy = cover %>% group_by(site_code, Taxon) %>%
  summarise(ps_path = unique(ps_path),
            local_lifespan = unique(local_lifespan),
            functional_group = unique(functional_group),
            N_fixer = unique(N_fixer))

site.covars <- read.csv("../Data/site_covars.csv")

site_pvals = read.csv("composition_pvalues.csv", header = TRUE, stringsAsFactors = FALSE)
site_pvals = site_pvals %>% rename(site.K = trt_K_num, site.N = trt_N_num, site.P = trt_P_num)

dot_full = read.csv("../data/dot_full.csv")

scalar1 <- function(x) {x / sqrt(sum(x^2))}
```

# Joining tables into master

```{r}
spec_dot = read.csv("../data/dot_full.csv")
head(spec_dot)
```

# Creating a "tuning table"

To assess how individual species contribute to community-patterns, we created a
"tuning table" that can be used to find covariates between species and their
estimated treatment responses, as well as subset our data to species that are
present in the first year of sampling, found in X number of plots over the course
of sampling, etc.

We generated this table using a set of potential variab

* What are the ways










```{r}
pretrt_total = ttable %>% group_by(site_code, Taxon, block) %>%
  summarise(pretrtcover = sum(pret_covN, pret_covControl, pret_covK, pret_covP)) %>%
  group_by(site_code) %>%
  summarise(pretrt_totalcover = sum(pretrtcover)) 
```

```{r}
pretrt = ttable %>% group_by(site_code, Taxon, block) %>%
  summarise(pretrtcover = sum(pret_covN, pret_covControl, pret_covK, pret_covP)) %>%
  group_by(site_code, Taxon) %>%
  summarise(pretrtcover = mean(pretrtcover)) %>%
  group_by(site_code) %>%
  mutate(relrank = rank(pretrtcover)/ max(rank(pretrtcover))) %>%
  mutate(species = gsub("\\.", " ", Taxon)) %>%
  select(-Taxon)
    
specscores = specscores_full %>%
  select(-X) %>%
  gather(key = "species",
         value = "coef", 
         -c(site, trt)) %>%
  mutate(species = gsub("\\.", " ", species)) %>%
  inner_join(pretrt, by = c(site = "site_code", species = "species"))

taxonomy = taxonomy %>% mutate(species = gsub("\\.", " ", Taxon)) %>%
  rename("site" = site_code)

specscores = specscores %>% plyr::join(taxonomy, by = c("site", "species"), match = "first")

specscores = specscores %>% inner_join(site_pvals %>% select(site, site.K, site.P, site.N))

specscores = specscores %>% inner_join(dot_full %>% select(site, NP, NK, PK))

resptable = specscores %>% inner_join(site.covars, by = c(site = "site_code"))

head(resptable)
```

```{r}
resptable %>% 
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>%
  filter(NP == min(NP)) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = pretrtcover, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE)
  
resptable %>% 
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>%
  filter(NP == max(NP)) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = pretrtcover, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE)

resptable %>% 
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>%
  filter(NP == median(NP)) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = pretrtcover, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE)
```

```{r}
resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>% 
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = pretrtcover, color = functional_group)) +
  facet_wrap(~functional_group) +
  stat_smooth(method = "lm", se = FALSE)


resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>% 
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = pretrtcover, color = ps_path)) +
  stat_smooth(method = "lm", se = FALSE)
```

```{r}
resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>% 
  filter(pretrtcover > .01) %>%
  ggplot(aes(x = functional_group, 
             y = trt_P,
             fill = functional_group)) +
  geom_boxplot()

resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>% 
  filter(pretrtcover > .01) %>%
  ggplot(aes(x = functional_group, 
             y = trt_N,
             fill = functional_group)) +
  geom_boxplot()

resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(site.P < .05 & site.N < .05) %>% 
  filter(pretrtcover > .01) %>%
  ggplot(aes(x = ps_path, 
             y = trt_P,
             fill = ps_path)) +
  geom_boxplot()

```

```{r}
formods = resptable %>%  
  spread(key = trt, value = coef) %>%
  filter(pretrtcover > .01) %>%
  filter(site.P < .05 & site.N < .05)

mod = summary(lm(trt_P ~ trt_N, data = formods))

summary(lm(trt_P ~ trt_N + functional_group, data = formods))

formods$resid = resid(mod)

formods %>%
  ggplot(aes(x = functional_group,
             y = resid,
             fill = functional_group)) +
  geom_boxplot()
```



```{r}
resptable %>% 
  filter(trt == "trt_N") %>%
  filter(functional_group %in% c("FORB", "GRAMINOID", "GRASS", "LEGUME", "WOODY")) %>%
  filter(pretrtcover > .01) %>%
  ggplot(aes(x = relrank, y = coef, color = functional_group)) + 
  geom_point() +
  stat_smooth(method = "lm")
```


```{r}
df = resptable %>% filter(trt_N < .05) %>% filter(trt == "trt_N")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```
```{r}
df = resptable %>% filter(trt_P < .05) %>% filter(trt == "trt_P")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```

```{r}
df = resptable %>% filter(trt_K < .05) %>% filter(trt == "trt_K")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```

# Fitting linear models to each piece of data?

```{r}
library(data.table)
compmat = read.csv("../data/compmat_edited.csv")
head(compmat)

longcomp <- compmat %>% select(-c("X")) %>%
  gather(key = "species",
         value = "cover", 
         -c("site_code", "year", "block", "plot", "trt", "year_trt", 
            "trt_K", "trt_P", "trt_N", "trt_K_num", "trt_P_num", "trt_N_num"))
data.table(longcomp)


fit_lm <- function(tmp){
  
  if(sum(cover) == 0){
    return()
  }else{
    fit = lmer(cover ~ trt_K_num + trt_P_num + trt_N_num + year_trt + (1|plot),
               data = tmp)
    
  }
  
}

tmp = longcomp[longcomp$site_code == "mcla.us",]
longcomp
```