---
title: "R Notebook"
output: html_notebook
---

LRR Niche Tradeoffs Approach 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(smatr)
library(tidyverse)
library(data.table)
library(dtplyr)
```

Compiled on `r date()`

```{r Calculating LRRs, include=FALSE}
# Read biomass data
biomass_data <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/comb-by-plot-09-April-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

biomass_data$site_code <- factor(biomass_data$site_code)
biomass_data$trt <- factor(biomass_data$trt)

#Choose sites with at least 3 years of data
sites <- biomass_data %>% 
  group_by(site_code,region,first_nutrient_year) %>% 
  summarise(yrs.data = length(unique(year)),fyear = min(year), lyear = max(year)) %>% 
  arrange(yrs.data)

sites.long <- as.character(sites[yrs.data >= 3,`site_code`])
rm(sites)

bmass <- biomass_data[site_code %in% sites.long]

cover <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/full-cover-09-April-2018.csv',
               stringsAsFactors = FALSE,
               na.strings = c('NA','NULL'))
cover.long <- cover[site_code%in%c(sites.long) & live==1,]
cover.long <- left_join(cover.long,distinct(bmass[,.(`site_code`,`region`)]))
```

```{r, include = FALSE}
#Choosing only our treatments of interest
cover.long <- cover.long[trt %in% c('Control','N','P','K','NPK','Fence'),]

#Clean out unknowns, non-vascular plants
sps <- unique(cover.long$Taxon)
unwanted.sps <- sps[c(grep('UNKNOWN',sps),grep('BRYOPHYTE',sps),grep('LICHEN',sps))]
#cover.long <- cover.long[!(Taxon %in% unwanted.sps | local_lifeform == 'MOSS' | functional_group == 'BRYOPHYTE')]
cover.long[functional_group=='GRAMINOID', functional_group := 'GRASS']


cover.long %>% filter(site_code == "hopl.us") %>% filter(year_trt == 0) %>% group_by(plot) %>% summarise(count = n())

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% 
  group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% 
  filter(yrs.data>=3)

## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% 
  select(-min.plot,-yrs.data)

cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)

cover.long$max_cover <- as.numeric(cover.long$max_cover)
```

Casting to wide-ish format data - allows for calculation of LRRs at a given year

What I think is problematic here is that even at year 0, there's a positive relationship between the log response ratio of N and the log response ratio of P -- if this just happened in a couple select sites, I don't think it's a problem, but this is a consistent pattern. If anything, we see that the slope of the line becomes increasingly flattened over time.

```{r}

trtcov <- cover.long %>% ungroup %>% select(-subplot) %>% select(-plot) %>% spread(key = trt, value = max_cover, fill=0)

trtcov = trtcov %>% mutate(LRR_N = log((N + 1) / (Control + 1)),
                  LRR_P = log((P + 1) / (Control + 1)),
                  LRR_K = log((K + 1) / (Control + 1)),
                  LRR_Fence = log((Fence + 1) / (Control + 1)),
                  LRR_NPK = log((K + 1) / (Control + 1)),
                  ABS_N = N - Control,
                  ABS_P = P - Control,
                  ABS_K = K - Control,
                  ABS_Fence = Fence - Control,
                  ABS_NPK = NPK - Control,
                  meancov = (N + P + K) / 3)

trtcov %>% filter(site_code == "cdcr.us") %>% 
  # filter(Control != 0 & P != 0 & N != 0) %>%
  ggplot(aes(x = LRR_N, y = LRR_P, color = Taxon)) + geom_point() +
  stat_smooth(method = "lm", aes(color = NULL), se = FALSE) + facet_wrap(~year_trt) +
  guides(color = "none")
```


# Subsetting to relevant species

```{r}
cover.long %>% group_by(site_code, block, Taxon) %>% summarise()

```


# cast long into wide
cover.wide <- dcast(cover.long,site_code+year+block+plot+trt ~ Taxon,
                    value.var='max_cover', 
                    fun.aggregate = sum,drop=T,fill=0)