---
title: "Nut Net 2018 Workshop"
output:
  html_document:
    df_print: paged
---

# Species-specific directional change

Given that we see that multiple nutrient addition produces greater change than just single nutrient additions,
do we expect that nutrient identity matters - does the addition of one limiting nutrient produce consistent changes in 
species-specific directional change.

What sites/blocks show consistent directional change?

* There should be more consistent directional change when there is a tradeoff with nutrient enrichment strategy
    * The more spatial heterogeneity in environmental heterogeneity, the more likelihood that different resource use strategies predominate
    * Temporal change and its relationship?

* What species are associated with directional change?

Steps to analyses:

```{r, message = FALSE, echo = FALSE, warning = FALSE}
################################################################################
### Data preprocessing

# Loading necessary packages
library(codyn) # Temporal analysis
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lubridate)
library(data.table)
library(vegetarian) # Exponentiated diversity indices (Hill Numbers)
library(lme4) # Linear mixed effects models
library(vegan)
library(data.table)
library(testthat)
library(plot3D)

################################################################################
# Loading in dataset (April 2018 Data)
cov.long <- fread('C:/Users/ebatz/Dropbox/NutNet Data/full-cover-09-April-2018.csv',
                  na.strings = c('NA','NULL'))

plot.descriptions <- read.csv("C:/Users/ebatz/Dropbox/NutNet Data/all-plot-descriptions-1-December-2017.csv",
                              stringsAsFactors = FALSE)

# Subsetting to just California sites
# site.subset <- c("mcla.us", "hopl.us", "sier.us", "elliot.us", "sedg.us",
#                  "cdcr.us", "konz.us", "cbgb.us", "trel.us", "cdpt.us")

# cov.long <- cov.long[cov.long$site_code %in% site.subset,]

# Choose identifying variables
ids <- c('site_code','year','year_trt', 'block','plot','trt')

# Remove non-live percent cover esimates
cov.long[,max_cover:=as.numeric(max_cover)] 
cov.long <- cov.long[live==1]

# Cast long into wide
plot.sp.mat <- dcast(cov.long,site_code+year+block+plot+trt+year_trt ~ Taxon,value.var='max_cover', 
                     fun.aggregate = sum,drop=T,fill=0)
```

# Subsetting by sites that have >= 5 years of data and removing all species with counts == 0

```{r}

# Selecting sites with years greater than or equal to 10
lt_sites = plot.sp.mat %>% 
  group_by(site_code) %>% 
  filter(!is.na(block)) %>%
  summarise(uniqueyears = length(unique(year))) %>%
  filter(uniqueyears >= 5) %>%
  select(site_code)

# Filtering based on those site codes
dat = plot.sp.mat %>% 
  filter(site_code %in% c(lt_sites$site_code)) %>% 
  filter(trt %in% c("N", "P", "K", "Control"))

# Selecting columns with all values greater than 0
dat = bind_cols(dat[,1:6], dat[,7:ncol(dat)][,colSums(dat[,7:ncol(dat)]) > 0])
dat = na.omit(dat)

# Binding these with dummy-variable versions of nutrient addition
com.attr <- dat[,1:6]
com.attr <-  inner_join(com.attr, plot.descriptions, by = c("site_code", "block", "plot"))
com.attr$N_num = as.numeric(com.attr$N) * com.attr$year_trt
com.attr$P_num = as.numeric(com.attr$P) * com.attr$year_trt
com.attr$K_num = as.numeric(com.attr$K) * com.attr$year_trt

# Pulling out community matrix
com.mat <- dat[,7:ncol(dat)]
expect_true(nrow(com.mat) == nrow(com.attr))
```

# Running permutational ANOVA

```{r}

# Creating list for storage
output = list()
# Initializing counter variables
counter <- 1

# Setting treatment as a factor (can use in other runs)
com.attr$trt <- factor(com.attr$trt, levels = c("Control", "N", "P", "K"))

# For all unique sites selected
for(i in unique(com.attr$site_code)){
  
  # Subset to a single site
  com.subset <- data.frame(com.mat)[com.attr$site_code == i,]
  
  # Remove all zero columns
  com.subset = com.subset[,colSums(com.subset) > 0]
  
  # Pull out relevant plot attributes for each site
  attr.subset <- data.frame(com.attr) %>% filter(site_code == i)
  
  # Check that these two matrices are the same size
  expect_true(nrow(attr.subset) == nrow(com.subset))
  
  # If there's only 1 block per site, add no block interactions
  if(length(unique(attr.subset$block)) == 1){ 
  mod_anov <- adonis(decostand(com.subset, method = "total") ~ trt * as.factor(year_trt) + as.factor(year_trt), 
                     #strata = as.factor(attr.subset$block), 
                      permutations = 99,
                      data = attr.subset,
                      method = "bray")
  }else{ # If there is more than one block, add block interactions
  mod_anov <- adonis(decostand(com.subset, method = "total") ~ trt * as.factor(year_trt)  +
                       trt * as.factor(block) + 
                       as.factor(block) * as.factor(year_trt), 
                     #strata = as.factor(attr.subset$block), 
                      permutations = 99,
                      data = attr.subset,
                      method = "bray")
  }
  
  # Saving ouput
  output[[counter]] = list(sitename = i,
                           aovtable = mod_anov$aov.tab,
                           sitescores = mod_anov$coef.sites,
                           specscores = mod_anov$coefficients)
  
  # Update counter
  counter = counter + 1
}

# Binds coefficients from each site together:
coefs = data.frame()
for(sites in 1:length(output)){
  if(sites == 1){
    coefs = data.frame(t(output[[sites]]$specscores))
    coefs$species = rownames(coefs)
    coefs$site = rep(output[[sites]]$sitename, nrow(coefs))
  }else{
    tojoin =  data.frame(t(output[[sites]]$specscores))
    tojoin$species = rownames(tojoin)
    tojoin$site = rep(output[[sites]]$sitename, nrow(tojoin))
    coefs = bind_rows(coefs, tojoin)
  }
}
```

# Plotting relationships of axis loadings
```{r}
coefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt1, 
             y = trt2,
         fill = site)) +
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  # stat_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  stat_smooth(method = "lm", aes(fill = NULL), fullrange = TRUE) +
  guides(fill = FALSE) +
  ggtitle("N / P Change") +   
  ylim(-.15, .15) +
  xlim(-.15, .15) +
  xlab("Nitrogen Effect") +
  ylab("Phosphorous Effect")
  #facet_wrap(~site)

ggsave("N-P Corr.jpeg", height = 7, width = 10)


coefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt1, 
             y = trt3,
         fill = site)) +
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  # stat_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  stat_smooth(method = "lm", aes(fill = NULL), fullrange = TRUE) +
  guides(fill = FALSE) +
  ggtitle("N / K Change") + 
  ylim(-.15, .15) +
  xlim(-.15, .15) +
  xlab("Nitrogen Effect") +
  ylab("Phosphorous Effect")

ggsave("N-K Corr.jpeg", height = 7, width = 10)

coefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt2, 
             y = trt3,
         fill = species)) +  
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  geom_smooth(method = "lm", fullrange = TRUE, aes(fill = NULL)) +
  guides(fill = FALSE) +
  ggtitle("K / P Change") +
  ylim(-.15, .15) +
  xlim(-.15, .15)
ggsave("K-P Corr.jpeg", height = 7, width = 10)


length(unique(data.frame(coefs %>%
  filter(X.Intercept. > .05) %>% select(species))$species)) / nrow(coefs %>%
  filter(X.Intercept. > .05))

output[[1]][c(1,2)]


```

```{r}
# Principal comoponents of proportional change:
newcoefs = coefs %>% filter(X.Intercept. > .05)

hist((newcoefs %>% group_by(site) %>% summarise(count = n()))$count, breaks = 6)

spec_pca = princomp(newcoefs[,c(2,3,4)])
plot(spec_pca)
(spec_pca$sdev[1] + spec_pca$sdev[2] ) / sum(spec_pca$sdev)  
biplot(spec_pca, choices = c(1,2), cex = .5)
?biplot
levels(com.attr$trt)
```

```{r}
# x, y, z variables
x <- coefs$trt1
y <- coefs$trt2
z <- coefs$trt3

# Compute the linear regression (z = ax + by + d)
fit <- loess(z ~ x + y, span = 1)

# predict values on regular xy grid
grid.lines = 26
x.pred <- seq(min(x), max(x), length.out = grid.lines)
y.pred <- seq(min(y), max(y), length.out = grid.lines)
xy <- expand.grid( x = x.pred, y = y.pred)
z.pred <- matrix(predict(fit, newdata = xy), 
                 nrow = grid.lines, ncol = grid.lines)

# fitted points for droplines to surface
fitpoints <- predict(fit)

# scatter plot with regression plane
jpeg(filename = "ResposneSurface.jpeg", height = 500, width = 800)

scatter3D(x, y, z, pch = 18, cex = .5, 
    theta = 45, phi = 45, ticktype = "detailed",
    xlab = "N", ylab = "P", zlab = "K",
    zlim = c(-.2, .2), ylim = c(-.2, .2), xlim = c(-.2, .2),
    surf = list(x = x.pred, y = y.pred, z = z.pred, facets = NA, fit = fitpoints), 
    main = "N-P-K Response Surface")

dev.off()
```

# What species to keep?

Must filter to only include the species that are present in:
* Most plots?
* All plots?
* Greater than X percent cover?

Switch to log-response ratio (detrended?)

```{r}
toselect = cov.long %>%
  filter(live == 1, trt %in% c("Control", "N", "P", "K")) %>%
  group_by(site_code, Taxon) %>%
  summarise(mcov = mean(max_cover),
            occ = n(),
            pres = paste0(unique(trt), collapse = "", sep = "_"),
            ntrts = length(unique(trt)))

selected = toselect %>% filter(ntrts == 4 & mcov > 5)

selected$Taxon <- gsub(" ", ".", selected$Taxon)

newcoefs = inner_join( coefs, selected, by = c("site" = "site_code", "species" = "Taxon"))


```



# Plotting relationships of axis loadings
```{r}

newcoefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt1, 
             y = trt2,
         fill = site)) +
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  # stat_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  stat_smooth(method = "lm", aes(fill = NULL), fullrange = TRUE) +
  guides(fill = FALSE) +
  ggtitle("N / P Change") +   
  ylim(-.15, .15) +
  xlim(-.15, .15) +
  xlab("Nitrogen Effect") +
  ylab("Phosphorous Effect")
  #facet_wrap(~site)

newcoefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt1, 
             y = trt3,
         fill = site)) +
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  # stat_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  stat_smooth(method = "lm", aes(fill = NULL), fullrange = TRUE) +
  guides(fill = FALSE) +
  ggtitle("N / K Change") + 
  ylim(-.15, .15) +
  xlim(-.15, .15) +
  xlab("Nitrogen Effect") +
  ylab("Phosphorous Effect")

newcoefs %>%
  filter(X.Intercept. > .05) %>%
  ggplot(aes(x = trt2, 
             y = trt3,
         fill = site)) +  
  geom_hline(yintercept = 0) + 
  geom_point(shape = 21, size = 3) +
  geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, aes(fill = NULL)) +
  guides(fill = FALSE) +
  ggtitle("K / P Change") +
  ylim(-.15, .15) +
  xlim(-.15, .15)
```
How to fit models?
* Treatment as numeric?
* Block as a separate factor (or as strata?) 
* Block interactions

```{r}
length(unique(newcoefs$site))
```



