---
title: "Observational Data Analysis"
output:
  prettydoc::html_pretty:
    highlight: github
---

Compiled on `r date()`

## Response:

For this initial analysis of obsevational data, I chose a quite restrictive subset of sites with:

* 30 plots total
* 3 blocks
* Pretreatment soil data on the three relevant nutrients (N, P, K)

Yielding 38 sites total for analysis. 

## Questions:

* Are measures of N, P, and K correlated with one another in sites? Or is variation of soil characteristics independent from one another?

* How well do soil variables explain within-site variation in community composition? What proportion of variance can be explained by soil nutrient Z-scores?

* Is the proportion of variance explained by soil nutrients dependent on variation in the soil resources themselves? I.e. is the influence of N, P, and K dependent on respective coefficients of variation of these resources?

* How correlated are community axis loadings of of N, P and K to one another? Are these relationships negative signalling some sort of resource use trade-off, or positive, suggesting that communities respond relatively uniformly to increased productivity?

## Methods:

* Subsetting species matrix and soil data to relevant sites
* Calculating Z-scores for soil nutrients
* Running canonical correspondence analysis (CCA) on site-wide community variance
* Extracting p-value of CCA, prop. variance explained, and predicted relationships between communities and environmnetal variables

## Highlights

* Weak correlations among soil resource (N, P, and K) Z-scores
* Substantial between-site differences in proportion variance explained by CCA
* Significant relationship between site-wide nitrogen coefficient of variation and CCA results.
* General positive pairwise correlations between SMAs of community relationships between N, P, and K.

## Analysis:

(Hidden in markdown output -- see source file)

```{r, include = FALSE}
################################################################################
### Data preprocessing

# Loading necessary packages
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(vegetarian) # Exponentiated diversity indices (Hill Numbers)
library(lme4) # Linear mixed effects models
library(vegan)
library(data.table)
library(testthat)

################################################################################
# Loading in dataset (April 2018 Data)
cov.long <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/full-cover-09-April-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

plot.descriptions <- read.csv("C:/Users/ebatzer/Dropbox/NutNet Data/all-plot-descriptions-1-December-2017.csv",
                              stringsAsFactors = FALSE)


soil.chars <- read.csv("C:/Users/ebatzer/Dropbox/NutNet Data/comb-by-plot-clim-soil-diversity-09-Apr-2018.csv",
                              stringsAsFactors = FALSE)

# Choose identifying variables
ids <- c('site_code','year','year_trt', 'block','plot','trt')

# Remove non-live percent cover esimates
cov.long[,max_cover:=as.numeric(max_cover)] 
cov.long <- cov.long[live==1]

# Cast long into wide
plot.sp.mat <- dcast(cov.long,site_code+year+block+plot+trt+year_trt ~ Taxon,value.var='max_cover', 
                     fun.aggregate = sum,drop=T,fill=0)

plot.sp.mat$block <- as.numeric(plot.sp.mat$block)
plot.sp.mat$plot <- as.numeric(plot.sp.mat$plot)
plot.sp.mat$year <- as.numeric(plot.sp.mat$year)
plot.sp.mat$year_trt <- as.numeric(plot.sp.mat$year_trt)

sites_toselect = plot.sp.mat %>% 
  filter(year_trt == 0) %>%
  group_by(site_code) %>% 
  summarise(num_block = length(unique(block)),
            num_plots = length(unique(plot))) %>%
  filter(num_block == 3 & num_plots == 30) %>%
  select(site_code)

sites_toselect = c(sites_toselect$site_code)

```

```{r, include = FALSE}
plot.sp.mat = plot.sp.mat %>% filter(site_code %in% sites_toselect)

# Subset by soil characteristics (must include pre-treatment data)
pretrt = soil.chars %>%   
  filter(year_trt == 0) %>%
  group_by(site_code, block) %>%
  summarise(ndat = sum(!is.na(pct_N))) %>%
  filter(ndat > 0)

# Number of sites with pretrt soil data:
length(unique(pretrt$site_code))

# Subsetting species matrix by these site codes
soil.chars = soil.chars[soil.chars$site_code %in% unique(pretrt$site_code), ] %>% filter(year_trt == 0)
plot.sp.mat = plot.sp.mat[plot.sp.mat$site_code %in% unique(pretrt$site_code), ]
plot.sp.mat = plot.sp.mat %>% filter(year_trt == 0)

soil.means = soil.chars %>% group_by(site_code) %>% 
  select(c("pct_C", "pct_N", "ppm_P", "ppm_K", "ppm_Ca", "ppm_Mg", "ppm_S", 
           "ppm_Na", "ppm_Zn", "ppm_Mn", "ppm_Fe", "ppm_Cu", "ppm_B", "pH")) %>%
  summarise_all(funs(mean = mean,
                     sd = sd))

soil.zscores = soil.chars %>% group_by(site_code, block, plot, trt, year, year_trt) %>% 
  select(c("plot", "pct_C", "pct_N", "ppm_P", "ppm_K", "ppm_Ca", "ppm_Mg", "ppm_S", 
           "ppm_Na", "ppm_Zn", "ppm_Mn", "ppm_Fe", "ppm_Cu", "ppm_B", "pH")) %>%
  left_join(soil.means, by = c("site_code")) %>%
  mutate(C_z = (pct_C - pct_C_mean) / pct_C_sd,
         N_z = (pct_N - pct_N_mean) / pct_N_sd,
         P_z = (ppm_P - ppm_P_mean) / ppm_P_sd,
         K_z = (ppm_K - ppm_K_mean) / ppm_K_sd,
         Ca_z = (ppm_Ca - ppm_Ca_mean) / ppm_Ca_sd,
         Mg_z = (ppm_Mg - ppm_Mg_mean) / ppm_Mg_sd,
         S_z = (ppm_S - ppm_S_mean) / ppm_S_sd,
         Na_z = (ppm_Na - ppm_Na_mean) / ppm_Na_sd,
         Zn_z = (ppm_Zn - ppm_Zn_mean) / ppm_Zn_sd,
         Mn_z = (ppm_Mn - ppm_Mn_mean) / ppm_Mn_sd,
         Fe_z = (ppm_Fe - ppm_Fe_mean) / ppm_Fe_sd,
         Cu_z = (ppm_Cu - ppm_Cu_mean) / ppm_Cu_sd,
         B_z = (ppm_B - ppm_B_mean) / ppm_B_sd,
         pH_z = (pH - pH_mean) / pH_sd
         )

alldat = inner_join(na.omit(soil.zscores), plot.sp.mat, by = c("site_code", "block", "plot", "year", "trt", "year_trt"))
alldat = bind_cols(alldat[,1:61], alldat[,62:ncol(alldat)][,colSums(alldat[,62:ncol(alldat)]) > 0])

# Subsetting data into two matrices
com.attr = alldat[,1:61]
com.mat = alldat[,62:ncol(alldat)]

# Checking dataframes
# head(com.mat)
# head(com.attr)
```

```{r, include = FALSE}
# For each site in the dataset:
  # Segment into blocks (if/if not)
  # Calculate site scores within each block
  # Indirect gradient analysis of N, P, and K on 
  # Pull out individual site scores within each block for each of the three nutrients (N, P, and K) 
  # Also pull out total variation explained by each axis (constrained vs not) + significance
  # Correlations between axes?

# What are the best ways to look at multi-block variation? 

# Alternatively, could do the same using CCA
output <- list()
counter <- 1

# For all unique sites selected
for(i in unique(com.attr$site_code)){
  
    # Subset to a single site
    com.subset <- com.mat[com.attr$site_code == i & com.attr$year_trt == 0,]
    
    # Remove all zero columns
    com.subset = com.subset[,colSums(com.subset) > 0]
  
    # Pull out relevant plot attributes for each site
    attr.subset <- data.frame(com.attr) %>% filter(site_code == i & year_trt == 0)
    
    # Check that these two matrices are the same size
    expect_true(nrow(attr.subset) == nrow(com.subset))

        
    # Standardizing community data matrix (normalized, in this case)
    mat.stdized <- decostand(com.subset, "total")
        
    # Running CCA on this data with the 3 axes of environmental variation (N, P, K)
    cca.mod = cca(mat.stdized ~ (N_z + P_z + K_z) , data = attr.subset)
      
    # Saving ouput
    output[[counter]] = list(sitename = i,
                               block = 1,
                               envscores = calibrate(cca.mod),
                               varexp = data.frame(mod = anova(cca.mod)$ChiSquare[1],
                                                   resid = anova(cca.mod)$ChiSquare[2],
                                                   dfmod = anova(cca.mod)$Df[1],
                                                   dfresid = anova(cca.mod)$Df[2],
                                                   pval = anova(cca.mod)$"Pr(>F)"[1]
                                                  ),
                               envvifs = vif.cca(cca.mod))
      
      # Update counter
      counter = counter + 1
    
}

```

### Are measures of N, P, and K correlated with one another?

For this step, I've simply calculated correlations between within-site Z-scores of N, P, and K within each plot. Z-scores of soil resources seem to be generally weakly correlated with one another within a given site, particularly with respect to P and K. Given this fact, I think our statistical power in detecting trade-offs may be a little limited.

```{r, echo = FALSE}
cat(paste("N-P Z-score Corr:",cor(com.attr$N_z, com.attr$P_z),
      "\nN-K Z-score Corr:", cor(com.attr$N_z, com.attr$K_z),
      "\nP-K Z-score Corr:", cor(com.attr$K_z, com.attr$P_z)))
```

### What sites exhibit the greatest proportion of community variance explained by soil nutrient Z-scores?

Here I've run a CCA (may have some favorable properties relative to RDA, but functions similarly) on each site, which should allow us to determine:

* What fraction of the overall variance in observational community composition we can attribute to variation in soil N, P, and K (and to test this fraction with ANOVA)
* What is the estimated relationship between community composition and soil characteristics (plot scores on N, P, and K axes)

Because I've subset our data to just those plots containing 30 plots in three blocks, these comparisons should be done with very similar datasets - similar to Jon Bakker's approach.

Here we see that a number of sites have significant fractions of community variance that can be explained by soil variables in CCA, and there's quite a range therein.

```{r, echo = FALSE}

# Binds coefficients from each site together:

for(sites in 1:length(output)){
  if(sites == 1){
    
    # Nutrient Loading Coefficients:
    coefs = data.frame(output[[sites]]$envscores)
    coefs$site = rep(output[[sites]]$sitename, nrow(coefs))
    coefs$block = rep(output[[sites]]$block, nrow(coefs))
    
    # P - values:
    pvals = data.frame(output[[sites]]$varexp)
    pvals$site = rep(output[[sites]]$sitename, nrow(pvals))
    pvals$block = rep(output[[sites]]$block, nrow(pvals))
    
  }else{
    
    # Nutrient Loading Coefficients
    tojoin = data.frame(output[[sites]]$envscores)
    tojoin$site = rep(output[[sites]]$sitename, nrow(tojoin))
    tojoin$block = rep(output[[sites]]$block, nrow(tojoin))
    coefs = bind_rows(coefs, tojoin)
    
    # P - values:
    tojoin = data.frame(output[[sites]]$varexp)
    tojoin$site = rep(output[[sites]]$sitename, nrow(tojoin))
    tojoin$block = rep(output[[sites]]$block, nrow(tojoin))
    pvals = bind_rows(pvals, tojoin)
  }
}

signif_sites <- pvals$site[pvals$pval <= 0.05]
pvals$varfrac = pvals$mod / (pvals$mod + pvals$resid)

pvals %>% arrange(varfrac) %>% ggplot(aes(x = site,
                     y = mod / (mod + resid),
                     color = pval <= .05)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### What sites show the greatest amount of variance attributable to soil resource variation?

Relating this variance fraction to CV's of N, P, and K shows an interesting, although not unexpected, pattern where the amoung of variance explained by soil nutrient Z-scores was correlated with the CV of nitrogen.

```{r, echo = FALSE}
p1 = left_join(pvals, soil.means, by = c("site" = "site_code")) %>%
  mutate(CV_n =  pct_N_sd / pct_N_mean,
         CV_p =  ppm_P_sd / ppm_P_mean,
         CV_k =  ppm_K_sd / ppm_K_mean) %>%
  ggplot(aes(x = CV_n,
             y = varfrac)) +
  geom_point(aes(color = site)) +
  guides(color = FALSE)+
  stat_smooth(method = "lm") +
  xlab("Coefficient of Variation - N") +
  ylab("Prop. Obs. Variance")

p2 = left_join(pvals, soil.means, by = c("site" = "site_code")) %>%
  mutate(CV_n =  pct_N_sd / pct_N_mean,
         CV_p =  ppm_P_sd / ppm_P_mean,
         CV_k =  ppm_K_sd / ppm_K_mean) %>%
  ggplot(aes(x = CV_p,
             y = varfrac)) +
  geom_point(aes(color = site)) +
  guides(color = FALSE)+
  stat_smooth(method = "lm") +
  xlab("Coefficient of Variation - P") +
  ylab("Prop. Obs. Variance")

p3 = left_join(pvals, soil.means, by = c("site" = "site_code")) %>%
  mutate(CV_n =  pct_N_sd / pct_N_mean,
         CV_p =  ppm_P_sd / ppm_P_mean,
         CV_k =  ppm_K_sd / ppm_K_mean) %>%
  ggplot(aes(x = CV_k,
             y = varfrac)) +
  geom_point(aes(color = site)) +
  guides(color = FALSE) +
  stat_smooth(method = "lm") +
  xlab("Coefficient of Variation - K") +
  ylab("Prop. Obs. Variance")

gridExtra::grid.arrange(p1, p2, p3, nrow = 2)

dat = left_join(pvals, soil.means, by = c("site" = "site_code")) %>%
  mutate(CV_n =  pct_N_sd / pct_N_mean,
         CV_p =  ppm_P_sd / ppm_P_mean,
         CV_k =  ppm_K_sd / ppm_K_mean)

car::Anova(lm(varfrac ~ CV_n * CV_p * CV_k, data = dat), type=2)

```

### How correlated are community axis loadings of of N, P and K to one another? Are these relationships negative signalling some sort of resource use trade-off, or positive, suggesting that communities respond relatively uniformly to increased productivity?

There's quite a bit of variation in the slope of these relationships. Some show very strong positive relationships, while others are more negative.

```{r, echo = FALSE }
p1 <- coefs %>% 
  ggplot(aes(x = N_z, y = P_z, color = site)) +
  geom_point(alpha = .1) +
  guides(color = FALSE)+
  stat_smooth(method = "lm", se = FALSE) +
  xlab("N Z-score Coefficient ") +
  ylab("P Z-score Coefficient")

p2 <- coefs %>% 
  ggplot(aes(x = N_z, y = K_z, color = site)) +
  geom_point(alpha = .1) +
  guides(color = FALSE)+
  stat_smooth(method = "lm", se = FALSE) +
  xlab("N Z-score Coefficient ") +
  ylab("K Z-score Coefficient")

p3 <- coefs %>% 
  ggplot(aes(x = P_z, y = K_z, color = site)) +
  geom_point(alpha = .1) +
  guides(color = FALSE)+
  stat_smooth(method = "lm", se = FALSE) +
  xlab("P Z-score Coefficient ") +
  ylab("K Z-score Coefficient")

gridExtra::grid.arrange(p1, p2, p3, nrow = 2)

cat(paste("N-P Corr:", cor(coefs$N_z, coefs$P_z),
      "\nN-K Corr:", cor(coefs$N_z, coefs$K_z),
      "\nP-K Corr:", cor(coefs$P_z, coefs$K_z)))
```

### Describing the above relationships with SMAs

I've repurposed your code here to explain these same relationships using SMAs. As before, significant relationships are in black, non-significant relationships are in grey. Generally, estimated community responses to N, P, and K are positive, though there may be a few negative relationships. Only one of these appears to be significant, and may be very outlier driven. Would be worth to check on these soil measurements. 

As an aside, the SMAs don't seem to ever assume a flat line. Is this a feature of the method? Or of the data itself?

```{r, echo = FALSE}
library(smatr)
modNP <- sma(N_z ~ P_z*site ,data=coefs, robust = TRUE)

sma.NP <- data.table(modNP$groupsummary) %>% select(-contains('test')) %>% rename(site=group) %>% mutate(slope.sig = (pval<0.05))

sma.NP <- left_join(sma.NP, coefs %>% group_by(site) %>% summarize(min.x = min(N_z,na.rm=T), max.x = max(P_z,na.rm=T))) %>% mutate(y.start = Slope * min.x + Int,y.end = Slope * max.x + Int)

ggplot(sma.NP,aes()) +
  geom_segment(aes(x=min.x,y=y.start,xend=max.x,yend=y.end,color=slope.sig)) +
  scale_color_manual(values=c('grey','black')) +
  geom_hline(yintercept=0,linetype='dotted') + 
  geom_vline(xintercept=0,linetype='dotted') + 
  theme_bw() + guides(color=F) +
  labs(x='N Axis Score',y='P Axis Score')
```
```{r, echo = FALSE}
modNK <- sma(N_z ~ K_z*site,data=coefs, robust = TRUE)

sma.NK <- data.table(modNK$groupsummary) %>% select(-contains('test')) %>% rename(site=group) %>% mutate(slope.sig = (pval<0.05))
sma.NK <- left_join(sma.NK, coefs %>% group_by(site) %>% summarize(min.x = min(N_z,na.rm=T), max.x = max(K_z,na.rm=T))) %>% mutate(y.start = Slope * min.x + Int,y.end = Slope * max.x + Int)

ggplot(sma.NK,aes()) +
  geom_segment(aes(x=min.x,y=y.start,xend=max.x,yend=y.end,color=slope.sig)) +
  scale_color_manual(values=c('grey','black')) +
  geom_hline(yintercept=0,linetype='dotted') + 
  geom_vline(xintercept=0,linetype='dotted') + 
  theme_bw() + guides(color=F) +
  labs(x='P Axis Score',y='K Axis Score')
```

```{r, echo = FALSE}
modPK <- sma(P_z ~ K_z*site,data=coefs, robust = TRUE)

sma.PK <- data.table(modPK$groupsummary) %>% select(-contains('test')) %>% rename(site=group) %>% mutate(slope.sig = (pval<0.05))
sma.PK <- left_join(sma.PK, coefs %>% group_by(site) %>% summarize(min.x = min(P_z,na.rm=T), max.x = max(K_z,na.rm=T))) %>% mutate(y.start = Slope * min.x + Int,y.end = Slope * max.x + Int)

ggplot(sma.PK,aes()) +
  geom_segment(aes(x=min.x,y=y.start,xend=max.x,yend=y.end,color=slope.sig)) +
  scale_color_manual(values=c('grey','black')) +
  geom_hline(yintercept=0,linetype='dotted') + 
  geom_vline(xintercept=0,linetype='dotted') + 
  theme_bw() + guides(color=F) +
  labs(x='N Axis Score',y='K Axis Score')
```