---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Individual Responses

To better understand why certain sites express correlations in their N, P, and K 
response trajectories, we can look at the individual responses of species that
constitute these community-level effects.

From existing work on nutrient enrichment (both in NutNet and other meta-analysis), 
there are a number of patterns that might be expected:

* Dominant species respond more positively to nutrient enrichment than non-dominants
* Certain functional groups (e.g. legumes) are likely to respond to P, K enrichment over others

These relationships may also give us more clues on what sort of site-level covariates
might be interested.

Here, I've taken the individual coefficients estimated from part 2 and plotted them
vs. a series of potential individual-level covariates. I've highlighted relationships
between dominance and functional group here, but a bunch more exploration is possible.

In the process of calculating these, I've produced a "tuning table" that has a number
of features that can also be used to subset the dataset prior to conducting these analyses:

* Pretrt presence / absence and cover
* Total observations over the course of sampling
* Pretrt dominance and control dominance (abundance / total observed cover)

```{r, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse);library(data.table);library(dtplyr); library(gridExtra); library(grid)

ttable <- fread("../Data/tuningtable.csv")

specscores_full <- read.csv("../Data/specscores.csv")

# taxonomy <- fread('C:/Users/ebatz/Dropbox/NutNet Data/Nutnet-site-taxonomy.csv',
#                stringsAsFactors = FALSE,
#                na.strings = c('NA','NULL'))

cover <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/full-cover-22-February-2019.csv')

taxonomy = cover %>% group_by(site_code, Taxon) %>%
  summarise(ps_path = unique(ps_path),
            local_lifespan = unique(local_lifespan),
            functional_group = unique(functional_group),
            N_fixer = unique(N_fixer))

site.covars <- read.csv("../Data/site_covars.csv")

site_pvals = read.csv("composition_pvalues.csv", header = TRUE, stringsAsFactors = FALSE)
site_pvals = site_pvals %>% rename(site.K = trt_K_num, site.N = trt_N_num, site.P = trt_P_num)

dot_full = read.csv("../data/dot_full.csv")

library(gridExtra)
```

```{r}
specchars = left_join(ttable, taxonomy) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon))

specscores_long = specscores_full %>%
  gather(key = "Taxon",
         value = "coef", 
         -c(site, trt)) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon)) %>%
  rename("site_code" = site)

full_responses = specscores_long %>% inner_join(specchars)
```

# Plotting relationships

Across all three fertilization treatment pairs (N-P, N-K, and P-K), responses
seem to be quite correlated. I've only showed correlations between the response
to N enrichment (trt_N) and P enrichment (trt_P):

```{r}
full_responses %>% 
  spread(key = trt, value = coef) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = dominanceControl, color = functional_group), alpha = .4) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("All Responses")
```

However, there is still a significant amount of site-to-site variation in the
shape of this line, equivalent to the correlations we examined in part 2. If we 
examine sites across the N-P tradeoff distribution, we might be able to 
get a sense of what species drive community-level trends.

#### Sites highlighted:

* Pinjarra Hills - Most negative NP response correlation

* Sagehen - Median NP response correlation

* Konza - Strongest NP response correlation

Some sites seem to show evidence that certain functional groups may be skewing these relationships -
Sagehen, for example, shows a legume species with high P affinity, but a weak positive response to N.

```{r}
full_responses %>% 
  spread(key = trt, value = coef) %>%
  filter(site_code == dot_full$site[dot_full$NP == min(dot_full$NP)]) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = dominanceControl, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("pinj.au")

full_responses %>% 
  spread(key = trt, value = coef) %>%
  filter(site_code == dot_full$site[dot_full$NP == median(dot_full$NP[-1])]) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = dominanceControl, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("sage.us")

full_responses %>% 
  spread(key = trt, value = coef) %>%
  filter(site_code == dot_full$site[dot_full$NP == max(dot_full$NP)]) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(size = dominanceControl, color = functional_group)) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("konz.us")
```

#### In a univariate perspective, there does not appear to be any significant correlation between functional group and response, however:

```{r, fig.height=10, fig.width=5}
p1 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .05) %>%
  ggplot(aes(x = functional_group, 
             y = trt_N,
             fill = functional_group)) +
  geom_boxplot()+
  ggtitle("N Response Coefficients")


p2 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .05) %>%
  ggplot(aes(x = functional_group, 
             y = trt_P,
             fill = functional_group)) +
  geom_boxplot() +
  ggtitle("P Response Coefficients")


p3 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .05) %>%
  ggplot(aes(x = functional_group, 
             y = trt_K,
             fill = functional_group)) +
  geom_boxplot() +
  ggtitle("K Response Coefficients")

grid.arrange(p1,p2,p3,
             top = textGrob("Dominance-Response Relationships",gp=gpar(fontsize=20,font=3)),
             nrow = 3)
```

#### Nor are there strong relationships with dominance (proportional of total cover in control plots over all years)

```{r, fig.height=10, fig.width=5}
p1 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .25) %>%
  ggplot(aes(x = dominanceControl, 
             y = trt_P,
             color = functional_group)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  ggtitle("N Response Coefficients")

p2 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .25) %>%
  ggplot(aes(x = dominanceControl, 
             y = trt_N,
             color = functional_group)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, fullrange = TRUE)+
  ggtitle("P Response Coefficients")

p3 = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(dominanceControl > .25) %>%
  ggplot(aes(x = dominanceControl, 
             y = trt_K,
             color = functional_group)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, fullrange = TRUE)+
  ggtitle("K Response Coefficients")


grid.arrange(p1,p2,p3,
             top = textGrob("Dominance-Response Relationships",gp=gpar(fontsize=20,font=3)),
             nrow = 3)
```

#### However, some additional exploration suggests that while species tend to share similar responses to all nutrient additions, functional groups exhibit differences in their deviations from this overall relationship.

Legumes, for example, seem to have a higher affinity to P than would be predicted by their +N treatment responses (same with K enrichment), while grasses have slightly stronger N affinity.

These results, calculated by examining residuals relative to the overall trend, are not shown graphically.


```{r, include = FALSE}
formodels = full_responses %>%  
  spread(key = trt, value = coef) %>%
  filter(functional_group %in% c("FORB", "GRAMINOID", "GRASS", "LEGUME"))

cat("Linear model of N-P relationship:\n")
summary(lm(trt_P ~ 0 + trt_N + functional_group, data = formodels))
cat("Linear model of N-K relationship:\n")
summary(lm(trt_K ~ 0 + trt_N + functional_group, data = formodels))
cat("Linear model of P-K relationship:\n")
summary(lm(trt_P ~ 0 + trt_K + functional_group, data = formodels))
```

# Next Steps

Before continuing to 






