---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Individual Responses

To better understand why certain sites express correlations in their N, P, and K 
response trajectories, we can look at the individual responses of species that
constitute these community-level effects.

From existing work on nutrient enrichment (both in NutNet and other meta-analysis), 
there are a number of patterns that might be expected:

* Dominant species respond more positively to nutrient enrichment than non-dominants
* Certain functional groups (e.g. legumes) are likely to respond to P, K enrichment over others

These relationships may also give us more clues on what sort of site-level covariates
might be interested.

Here, I've taken the individual coefficients estimated from part 2 and plotted them
vs. a series of potential individual-level covariates. I've highlighted relationships
between dominance and functional group here, but a bunch more exploration is possible.

In the process of calculating these, I've produced a "tuning table" that has a number
of features that can also be used to subset the dataset prior to conducting these analyses:

* Pretrt presence / absence and cover
* Total observations over the course of sampling
* Pretrt dominance and control dominance (abundance / total observed cover)

```{r, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse);library(data.table);library(dtplyr); library(gridExtra); library(grid)

ttable <- fread("../Data/tuningtable.csv")

specscores_full <- read.csv("../Data/tradeoffs_specscores.csv")

specscores_long <- specscores_full %>% 
  select(-X) %>% 
  gather(key = "Taxon", value = "Coef", - site, - trt) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon)) %>%
  mutate(Taxon = str_trim(Taxon)) %>%
  rename("site_code" = "site")

# taxonomy <- fread('C:/Users/ebatz/Dropbox/NutNet Data/Nutnet-site-taxonomy.csv',
#                stringsAsFactors = FALSE,
#                na.strings = c('NA','NULL'))

cover <- fread('../Data/full-cover-02-August-2019.csv')

taxonomy <- cover %>% group_by(site_code, Taxon) %>%
  filter(!is.null(Family)) %>%
  summarise(ps_path = unique(ps_path),
            functional_group = unique(functional_group),
            N_fixer = unique(N_fixer)) %>%
  mutate(functional_group = case_when(functional_group == "GRAMINOID" ~ "Graminoid",
                                      functional_group == "GRASS" ~ "Graminoid",
                                      functional_group == "LEGUME" ~ "Legume",
                                      functional_group == "WOODY" ~ "Woody",
                                      functional_group == "FORB" ~ "Forb")) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon)) %>%
  mutate(Taxon = str_trim(Taxon))

#site.covars <- read.csv("../Data/site_covars.csv")

site_pvals = read.csv("../Data/tradeoffs_RRPP_Pvals.csv", header = TRUE, stringsAsFactors = FALSE)
site_pvals = site_pvals %>% rename(site.K = trt_K_num, site.N = trt_N_num, site.P = trt_P_num)

dot_full = read.csv("../Data/dot_full.csv")

library(gridExtra)
```

```{r}
full_responses = specscores_long %>% inner_join(taxonomy) %>% filter(!is.na(functional_group))
```

# Plotting relationships

Across all three fertilization treatment pairs (N-P, N-K, and P-K), responses
seem to be quite correlated. I've only showed correlations between the response
to N enrichment (trt_N) and P enrichment (trt_P):

```{r}
full_responses %>% 
  spread(key = trt, value = Coef) %>%
  ggplot(aes(x = trt_N, y = trt_P)) +
  geom_point(aes(color = functional_group), alpha = .4) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("All Responses") +
  xlab("Beta_N") +
  ylab("Beta_P")

full_responses %>% 
  spread(key = trt, value = Coef) %>%
  ggplot(aes(x = trt_N, y = trt_K)) +
  geom_point(aes(color = functional_group), alpha = .4) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("All Responses") +
  xlab("Beta_N") +
  ylab("Beta_K")

full_responses %>% 
  spread(key = trt, value = Coef) %>%
  ggplot(aes(x = trt_P, y = trt_K)) +
  geom_point(aes(color = functional_group), alpha = .4) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("All Responses") +
  xlab("Beta_P") +
  ylab("Beta_K")
```

# Fitting with SMAs

```{r}
library(smatr)

smadat <- full_responses %>%
  spread(key = "trt", value = "Coef")

smafit <- sma(trt_P ~ trt_N * functional_group, smadat)
summary(smafit)

smafit$groupsummary %>%
  ggplot(aes(color = functional_group,
             slope = Slope,
             intercept = Int)) +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group), 
             data = smafit$groupsummary,
             size = 2) +
  xlim(-25, 25) +
  ylim(-25, 25) +
  xlab("N Response") +
  ylab("P Response")

smafit <- sma(trt_K ~ trt_N * functional_group, smadat)

smafit$groupsummary %>%
  ggplot(aes(color = functional_group,
             slope = Slope,
             intercept = Int)) +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group), 
             data = smafit$groupsummary,
             size = 2) +
  xlim(-25, 25) +
  ylim(-25, 25) +
  xlab("N Response") +
  ylab("K Response")

smafit <- sma(trt_K ~ trt_P * functional_group, smadat)
summary(smafit)

smafit$groupsummary %>%
  ggplot(aes(color = functional_group,
             slope = Slope,
             intercept = Int)) +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group), 
             data = smafit$groupsummary,
             size = 2) +
  xlim(-25, 25) +
  ylim(-25, 25) +
  xlab("P Response") +
  ylab("K Response")
```

# Conducting a PCA

```{r}
pc <- prcomp(smadat %>% select(trt_N, trt_K, trt_P))
pc$sdev[1] / sum(pc$sdev)
pcsummary <- summary(pc)$importance[2,]
data.frame(pcsummary) %>% 
  ggplot(aes(x = c("PC1", "PC2", "PC3"),
             y = pcsummary)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  ylab("Proportion of Variance") +
  xlab("Principal Component") +
  ggtitle("PC Decomposition")

ggsave("../Figures/tradeoffs_pcplot.jpeg", height = 6, width = 4)
```

# Fitting a 3d figure

```{r}
library(plot3D)
pc1_loadings <- pc$rotation[,1]
pc1_fit <- data.frame(t((pc1_loadings) %*% t(c(-25, 20))))
colnames(pc1_fit) = c("N", "P", "K")

theta = 30
phi = 30

# jpeg("../Figures/3dviz.jpeg", height = 5, width = 6)

scatter3D(z = smadat$trt_N, x = smadat$trt_P, y = smadat$trt_K, pch = 19, cex = 1, 
      theta = theta, phi = phi, col = "blue",
      xlab = "P Response", ylab = "K Response", zlab = "N Response", 
      main = "Overall Treatment Responses", alpha = .25, colkey = FALSE, ticktype = "detailed")

scatter3D(z = pc1_fit$N, x = pc1_fit$P, y = pc1_fit$K, pch = 19, lwd = 2, 
      theta = theta, phi = phi, col = "black", add = TRUE, type = "l", lty = 2)
```




