---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(smatr)
library(tidyverse)
library(data.table); library(dtplyr)
```

Compiled on `r date()`

```{r Calculating LRRs, include=FALSE}
rm(list=ls())

# Read biomass data
biomass_data <- fread('C:/Users/ebatz/Dropbox/NutNet Data/comb-by-plot-09-April-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

biomass_data$site_code <- factor(biomass_data$site_code)
biomass_data$trt <- factor(biomass_data$trt)

#Choose sites with at least 3 years of data
sites <- biomass_data %>% 
  group_by(site_code,region,first_nutrient_year) %>% 
  summarise(yrs.data = length(unique(year)),fyear = min(year), lyear = max(year)) %>% 
  arrange(yrs.data)

sites.long <- as.character(sites[yrs.data >= 3,`site_code`])
rm(sites)

bmass <- biomass_data[site_code %in% sites.long]

cover <- fread('C:/Users/ebatz/Dropbox/NutNet Data/full-cover-09-April-2018.csv',
               stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))
cover.long <- cover[site_code%in%c(sites.long) & live==1,]
cover.long <- left_join(cover.long,distinct(bmass[,.(`site_code`,`region`)]))
cover.long$max_cover = as.numeric(cover.long$max_cover)
```

```{r, include = FALSE}
#Choosing only our treatments of interest
cover.long <- cover.long[trt %in% c('Control','N','P','K','NPK','Fence')]

#Clean out unknowns, non-vascular plants
sps <- unique(cover.long$Taxon)
unwanted.sps <- sps[c(grep('UNKNOWN',sps),grep('BRYOPHYTE',sps),grep('LICHEN',sps))]
cover.long <- cover.long[!(Taxon %in% unwanted.sps | local_lifeform == 'MOSS' | functional_group == 'BRYOPHYTE')]
cover.long[functional_group=='GRAMINOID', functional_group := 'GRASS']
rm(sps,unwanted.sps)

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% 
  group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% 
  filter(yrs.data>=3)

## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% 
  select(-min.plot,-yrs.data)

cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)

# cast long into wide
cover.wide <- dcast(cover.long,site_code+year+block+plot+trt+year_trt ~ Taxon,
                    value.var='max_cover', 
                    fun.aggregate = sum,drop=T,fill=0)
```


```{r, include = FALSE}
# Adding in additional info

siteinfo <- biomass_data[site_code %in% sites.long, 
                         .(`site_code`,`site_name`,`continent`,`region`,`first_nutrient_year`,`site_richness`)] %>% distinct()

spsinfo <- select(data.frame(cover.long),site_code,Family:ps_path,-live) %>% distinct()

#Summarizing some information about the species
#Max cover reached in a plot in that site, CV of cover
# coeflist.site <- left_join(coeflist.site,cover.long %>% group_by(site_code, Family, Taxon) %>% summarize(max.cover = max(max_cover), cv.cover=sd(max_cover)/mean(max_cover)))
```

Standardizing by total observations per column - it may make sense to not just look at absolute cover, but rather, the % cover of each observed plot

```{r}
cols <- c(colnames(cover.wide)[-c(1:6)])
cover.wide[ , (cols) := lapply(.SD, "as.numeric"), .SDcols = cols]
cover.atts = cover.wide[,1:6]
cover.vals = data.frame(cover.wide[,-c(1:6)])
cover.stdize = round(vegan::decostand(cover.vals, method = "total"),2)
cover.mat = bind_cols(cover.atts, cover.stdize)
cover.mat[1:10, 1:10]
```

What species to keep? Ideally, we would have species that are:

- Present in all treatments (unlikely), or at least present in all treatments to start
- Persistent through time

```{r}
dat <- cover.long[,iden := paste0(site_code,Taxon,block)]
dat <- dat[iden %in% unique(obs.keep$iden)]
obs.keep[,iden:= NULL]
#convert to wide, but only at the site-block scale, to avoid too many zeros
# Add minimum observed percent cover (0.01) to all values - to the wide table
# Convert back to long
dat2 <- data.table()
#i='cdcr.us' ; j = 1
for (i in unique(dat$site_code)) {
  subdat <- dat[site_code==i]
  for (j in unique(subdat$block)) {
    subdat2 <- subdat[block==j]
    dat2.wide <- dcast(subdat2,site_code+year+year_trt+block+plot+trt ~ Taxon,value.var='max_cover', fun.aggregate = sum,drop=T,fill=0)
    #Now add 0.01 to all values
    dat2.wide[,7:ncol(dat2.wide)] <- dat2.wide[,7:ncol(dat2.wide)] + 0.01
    dat2.long <- melt(dat2.wide,id.vars=c(1:6),variable.name='Taxon',value.name='max_cover')
    dat2 <- bind_rows(dat2,dat2.long)
  }
}

rm(dat2.long,dat2.wide,i,j,subdat,subdat2)


```



I am concerned that detrending in this fashion is predisposed to constructing the 
