---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

# Filtering data with different thresholds:

Per meting on 9-6-2018, filtering the community dataset to a subset of species 
may greatly improve the relevance + significance of our result.

Key "levers" to manipulate:

* Normalization of cover values - dividing matrix values by row total
    * Maximum (total) cover = 1
* Mean cover in control
    * Mean cover in control plot over the observed sampling period
* Persistence in control
    * Number of observations of a species in control over the observed sampling period
* Present in year 0 for control and at least one nutrient treatment
    * Observations in initial year that may influence detrending

And correlations between between.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(smatr); library(tidyverse); library(data.table); library(dtplyr)
library(testthat); library(gridExtra)

# Read biomass data
biomass_data <- fread('C:/Users/ebatz/Dropbox/NutNet Data/comb-by-plot-31-August-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

biomass_data$site_code <- factor(biomass_data$site_code)
biomass_data$trt <- factor(biomass_data$trt)

#Choose sites with at least 3 years of data
sites <- biomass_data %>% 
  group_by(site_code,region,first_nutrient_year) %>% 
  summarise(yrs.data = length(unique(year)),fyear = min(year), lyear = max(year)) %>% 
  arrange(yrs.data)

sites.long <- as.character(sites[yrs.data >= 3,`site_code`])
rm(sites)

bmass <- biomass_data[site_code %in% sites.long]

cover <- fread('C:/Users/ebatz/Dropbox/NutNet Data/full-cover-31-August-2018.csv',
               stringsAsFactors = FALSE,
               na.strings = c('NA','NULL'))
cover.long <- cover[site_code%in%c(sites.long) & live==1,]
cover.long <- left_join(cover.long,distinct(bmass[,.(`site_code`,`region`)]))
cover.long <- cover.long[ , max_cover := as.numeric(max_cover)]

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% 
  group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% 
  filter(yrs.data>=3)

## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% 
  select(-min.plot,-yrs.data)

cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)
```

# Tuning function:

The tuning function produces a new datatable (tuningtable) that contains information
on the relevant variables we discussed selecting species by. It's a modular function,
so if we need to add more variables or treatments in our subsetting, it's easy to do so.

I think the easiest way to use this table is to select a number of thresholds,
filter the table by these thresholds, then use a left_join to filter the 
LRR data to just those relevant species.

Some of the key variables to pull from this dataset are:

* Persistence in control - the total number of observations in control divided 
by the number of years

* Mean cover in control - the average recorded cover in control treatments over 
the duration of sampling, including years with no observations (given a "0")

* Representation in the first year of sampling - the total number of plots where
a species was observed in year 0, pre-treatment.

The first few rows of the table are as follows:

```{r, include = FALSE, eval = FALSE}
tuneR <- function(covdata, # Cover data
                  selected.trts = c("Control", "N", "P", "K", "Fence", "NPK", "NPK + Fence"), # Treatments to include
                  normalize = TRUE, # Normalize or not? Logical
                  sigdig = 2 # Number of significant digits for rounding
){
  
  # If deciding to normalize data
  if(normalize == TRUE){
    
      # For all site x block x plot x year_trt combinations
    coversums <- covdata %>% group_by(site_code, block, plot, year_trt) %>%
      
      # Create a new column of total cover and join to dataset
      summarise(plottotal = sum(na.omit(max_cover)))

    # Setting new cover data with normalized cover
    covdata <- left_join(covdata, coversums) %>%
      
      # Reassign max_cover as normalized cover value
      mutate(max_cover = round(max_cover / plottotal, sigdig)) %>%
      
      # Remove the total cover value
      select(-plottotal)
    
    if(max(covdata$max_cover) > 1){
      stop(paste("Error - Max cover value not less than one in sites",
                 unique(covdata$site_code[covdata$max_cover > 1])))
    }
  }
  
  # General observation table
  obs.per.plot = data.frame(covdata) %>%
    filter(trt %in% selected.trts) %>%
    group_by(site_code, block) %>%
    summarise(nyears = length(unique(year)),
              firstyear = min(year),
              lastyear = max(year))
  
  # Frequency of species observations and mean cover
  summarise_obs <- function(x){
    
    # Create storage dataframe with total number of observations
    output = data.frame(totalobs = nrow(x))
    counter = 1
    
    # For each set of treatments within each block x species combination
    for(i in unique(selected.trts)){
      
      counter = counter + 1
      
      # Create and rename a new column that indicates the total times a species was found within a treatment
      output = bind_cols(output, new = sum(x$trt == i))
      colnames(output)[counter] = paste("tot_obs", i, sep = "")
      
      # And do the same for mean cover
      counter = counter + 1
      output = bind_cols(output, new = mean(na.omit(x$max_cover[x$trt == i])))
      colnames(output)[counter] = paste("mean_cov", i, sep = "")
    }
    
    return(output)
  }
  
  # Run function for all years
  specobs = data.frame(covdata) %>%
      filter(trt %in% selected.trts) %>%
      group_by(site_code, block, Taxon) %>%
      do(summarise_obs(.))
  
  # Presence in the first year of sampling
  summarise_yearzero <- function(x){
    
    # Create storage dataframe with total number of observations
    output = data.frame(total_pret_obs = nrow(x))
    counter = 1
    
# For each set of treatments within each block x species combination
    for(i in unique(selected.trts)){
      
      # Create and rename a new column that indicates whether a species was found within a treatment for the first year of sampling
      counter = counter + 1
      output = bind_cols(output, new = sum(x$trt == i))
      colnames(output)[counter] = paste("pret_obs", i, sep = "")
      
      # And do the same for cover
      counter = counter + 1
      newval = x$max_cover[x$trt == i]
      
      # If there is no recorded value, give it a zero
      if(length(newval) == 0){newval = 0}
      
      # And bind together as before
      output = bind_cols(output, new = newval)
      colnames(output)[counter] = paste("pret_cov", i, sep = "")
      }
    
    return(output)
  }
  
  # Run function for the first year
  trtcov <- data.frame(covdata) %>% 
    filter(trt %in% selected.trts) %>% 
    filter(year_trt == 0) %>% 
    group_by(site_code, block, Taxon) %>%
    do(summarise_yearzero(.))

  # Binding all values together into a single dataset
  fulloutput = inner_join(obs.per.plot, specobs, by = c("site_code", "block"))
  fulloutput = left_join(fulloutput, trtcov, by = c("site_code", "block", "Taxon"))
  fulloutput = fulloutput %>% replace(., is.na(.), 0) %>% ungroup()
  
  return(fulloutput)
}

tuningtable <- tuneR(cover.long)
write.csv(x = tuningtable, "../data/tuningtable_normalized.csv")
```

```{r}
tuningtable = read.csv("../data/tuningtable_normalized.csv", 
                       header = TRUE,
                       stringsAsFactors = FALSE)
head(tuningtable)
```


# Column Names in TuningTable

Site_code, block, plot as before

New columns:

__Plot level__  

* nyears - Number of years of observations (plot)                

* firstyear - First year of observation (plot)  

* lastyear - Last year of observation (plot)    

__Taxon level__  

* Taxon - Taxon name                

* totalobs - Total number of observations in block, over time   

* tot_obsControl - Total number of observations in control plot  

* mean_covControl - Mean cover in control plot, all years   

* tot_obsN - Total number of observations in +N plot  

* mean_covN - Mean cover in +N plot, all years   

* tot_obsP - Total number of observations in +P plot  

* mean_covP - Mean cover in control +P, all years   

* tot_obsK - Total number of observations in +K plot  

* mean_covK - Mean cover in +K plot, all years   

* tot_obsFence - Total number of observations in +Fence plot  

* mean_covFence - Mean cover in +Fence plot, all years   

* tot_obsNPK - Total number of observations in +NPK plot  

* mean_covNPK -  Mean cover in +NPK plot, all years   

* tot_obsNPK + Fence - Total number of observations in +NPK + Fence plot  

* mean_covNPK + Fence - Mean cover in +NPK + Fence plot, all years   

__Pretreatment, Taxon level__  

* total_pret_obs  - Total number of observations in block, first year  

* pret_obsControl - Observation in control plot, first year (1/0)  

* pret_covControl - Cover in control plot, first year  

* pret_obsN - Observation in +N plot, first year (1/0)  

* pret_covN - Cover in +N plot, first year   

* pret_obsP - Observation in +P plot, first year (1/0)  

* pret_covP - Cover in +P plot, first year  

* pret_obsK - Observation in +K plot, first year (1/0)  

* pret_covK - Cover in +K plot, first year  

* pret_obsFence - Observation in +Fence plot, first year (1/0)  

* pret_covFence - Cover in +Fence plot, first year  

* pret_obsNPK - Observation in +NPK plot, first year (1/0)  

* pret_covNPK - Cover in +NPK plot, first year   

* pret_obsNPK + Fence - Observation in +NPK + Fence plot, first year (1/0)   

* pret_covNPK + Fence - Cover in +NPK + Fence plot, first year  

__As an aside, some of the sites don't seem to be behaving well:__ 

The unique years recorded does not always seem to equal the first year subtracted
from the last year, + 1. Is this the result of years where no live species are 
recorded in a plot?

__Offending sites:__
```{r, echo = FALSE}
as.character(unique(tuningtable$site_code[tuningtable$lastyear - tuningtable$firstyear + 1 > tuningtable$nyears]))
```

# Correlations among measures:

```{r}
badsites = as.character(unique(tuningtable$site_code[tuningtable$totalobs > tuningtable$nyears * 7]))
tuningtable = tuningtable %>% filter(! site_code %in% badsites)

# Mean control cover
tuningtable %>% ggplot(aes(x = mean_covControl * (tot_obsControl / (tuningtable$lastyear - tuningtable$firstyear + 1)),
                           y = tot_obsControl / (tuningtable$lastyear - tuningtable$firstyear + 1))) +
  geom_point() +
  stat_smooth(se = FALSE) +
  ylab("Persistence in Control") +
  xlab("Mean Cover in Control") +
  ggtitle("Cover vs. Persistence")

# Persistence in control
tuningtable %>% filter(pret_obsControl == 1) %>%
  ggplot(aes(x = mean_covControl * (tot_obsControl / (lastyear - firstyear + 1)),
                           y = pret_obsControl + pret_obsN + pret_obsK + pret_obsP,
                           color = paste(pret_obsControl, pret_obsN, pret_obsK, pret_obsP))) +
  geom_point() +
  stat_smooth(se = FALSE, aes(color = NULL)) +
  ylab("Representation in Year 0") +
  xlab("Mean Cover in Control") +
  ggtitle("Cover vs. Representation in Year 0") +
  guides(color=guide_legend(title="Yr1 Presence \n     C|N|K|P"))


# Presence in year 0 + at least 1 nutrient addition treatment
tuningtable %>% filter(pret_obsControl == 1) %>%
  ggplot(aes(x = tot_obsControl / (lastyear - firstyear + 1),
             y = pret_obsControl + pret_obsN + pret_obsK + pret_obsP,
             color = paste(pret_obsControl, pret_obsN, pret_obsK, pret_obsP))) +
  geom_point() +
  stat_smooth(se = FALSE, aes(color = NULL)) +
  ylab("Representation in Year 0") +
  xlab("Persistence in Control") +
  ggtitle("Persistence vs. Representation in Year 0") +
  guides(color=guide_legend(title="Yr1 Presence \n     C|N|K|P"))
```

# LRR Sensitivity to Filtering Approaches

To see how the use of these different tuning parameters, I've constructed a series
of figures that subset your LRR output based on the tuning table. It's a bit messy,
but provides some good indication of threshold values for filtering.

Because many of the filtering variables are correlated, the results are quite 
similar as we get more selective.

```{r, warning = FALSE, message = FALSE}
# Normalizing cover data:
# For all site x block x plot x year_trt combinations
coversums <- cover.long %>% group_by(site_code, block, plot, year_trt) %>%
  
  # Create a new column of total cover and join to dataset
  summarise(plottotal = sum(na.omit(max_cover)))

# Setting new cover data with normalized cover
cover.long <- left_join(cover.long, coversums) %>%
  
  # Reassign max_cover as normalized cover value
  mutate(max_cover = round(max_cover / plottotal, 2)) %>%
  
  # Remove the total cover value
  select(-plottotal)

if(max(cover.long$max_cover) > 1){
  stop(paste("Error - Max cover value not less than one in sites",
             unique(cover.long$site_code[cover.long$max_cover > 1])))
}

#Clean out unknowns, non-vascular plants
sps <- unique(cover.long$Taxon)
unwanted.sps <- sps[c(grep('UNKNOWN',sps),grep('BRYOPHYTE',sps),grep('LICHEN',sps))]
cover.long <- cover.long[!(Taxon %in% unwanted.sps | local_lifeform == 'MOSS' | functional_group == 'BRYOPHYTE')]
cover.long[functional_group=='GRAMINOID',functional_group := 'GRASS']
rm(sps,unwanted.sps)

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% filter(yrs.data>=3)
## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% select(-min.plot,-yrs.data)
cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)

# cast long into wide
cover.wide <- dcast(cover.long,site_code+year+block+plot+trt ~ Taxon,value.var='max_cover', fun.aggregate = sum,drop=T,fill=0)


# Adding in additional info ----------------------------------------------------

siteinfo <- biomass_data[site_code %in% sites.long, .(`site_code`,`site_name`,`continent`,`region`,`first_nutrient_year`,`site_richness`)] %>% distinct()

spsinfo <- select(data.frame(cover.long),site_code,Family:ps_path,-live) %>% distinct()

#Summarizing some information about the species
#Max cover reached in a plot in that site, CV of cover
# coeflist.site <- left_join(coeflist.site,cover.long %>% group_by(site_code, Family, Taxon) %>% summarize(max.cover = max(max_cover), cv.cover=sd(max_cover)/mean(max_cover)))

# Independent measure of N,P and K response following EL 2013 ----------------------------------

# Intermediate data products
### dat2 - cover data from which LRRs are being calculated
### dat3 - baseline year data for each species-plot
### dat4 - LRR for each species plot relative to baseline
### dat5 - LRR of nut effect after detrending control

#Species selection - 	"only those species recorded at least once either pre or 
# post treatment in the nutrient treatment plot within a block"
# keep a species-site-block-trt only if has an observation in the Nut treated plots, at any point of time
obs.keep <- cover.long[trt != 'Control'] %>% group_by(site_code,Taxon,block,trt) %>% 
  summarize(num.yrs = length(unique(year))) %>% mutate(iden = paste0(site_code,Taxon,block))
# Use at the end of the procedure to only select the ones we want

dat <- cover.long[,iden := paste0(site_code,Taxon,block)]
dat <- dat[iden %in% unique(obs.keep$iden)]
obs.keep[,iden:= NULL]
#convert to wide, but only at the site-block scale, to avoid too many zeros
# Add minimum observed percent cover (0.01) to all values - to the wide table
# Convert back to long
dat2 <- data.table()
#i='cdcr.us' ; j = 1
for (i in unique(dat$site_code)) {
  subdat <- dat[site_code==i]
  for (j in unique(subdat$block)) {
    subdat2 <- subdat[block==j]
    dat2.wide <- dcast(subdat2,site_code+year+year_trt+block+plot+trt ~ Taxon,value.var='max_cover', fun.aggregate = sum,drop=T,fill=0)
    #Now add 0.01 to all values
    dat2.wide[,7:ncol(dat2.wide)] <- dat2.wide[,7:ncol(dat2.wide)] + .01
    dat2.long <- melt(dat2.wide,id.vars=c(1:6),variable.name='Taxon',value.name='max_cover')
    dat2 <- bind_rows(dat2,dat2.long)
  }
}

rm(dat2.long,dat2.wide,i,j,subdat,subdat2)


# Calculate LRR of each year relative to pretreatment year ----------------

#add in cover of pretrt year for each species-plot combination
dat3 <- dat2 %>% group_by(site_code,plot,Taxon) %>% filter(year_trt == min(year_trt)) 
dat3 <- dat3[,.(`site_code`,`plot`,`Taxon`,`year_trt`,`max_cover`)] %>% rename(base_year=year_trt,cover_base=max_cover)
#dat3 <- dat3[,.(`site_code`,`plot`,`Taxon`,`base_year`=year_trt,`cover_base`=max_cover)]
#    select(site_code,plot,Taxon,`base_year`=year_trt,`cover_base`=max_cover)
# sier.us, sgs.us and hopl.us show some species as having their min(year_trt) measurement in y1
# saline.us and saana.fi are missing pretrt year in the cover dataset
#table(dat1[site_code=='hopl.us']$plot,dat1[site_code=='hopl.us']$year)
dat4 <- left_join(dat2,dat3)
dat4 <- dat4[year_trt != base_year] #remove the base year
#dat4[,LRRt := log2(max_cover/cover_base)] ;
dat4[,LRRt := log2(max_cover/cover_base)] ;
#Chose log2 in order for the coefficients to be interpretable


# Detrend from the control ------------------------------------------------

#add a column of control for each species-block in a year
dat5 <- left_join(dat4[trt!='Control'],
                  select(dat4[trt=='Control'],site_code,Taxon,year,year_trt,block,plot_cont=plot,cover_cont=max_cover,cover.base_cont=cover_base,LRRcont_yt=LRRt))
dat5 <- rename(dat5,LRRnut_yt=LRRt,plot_nut=plot,cover_nut=max_cover,cover.base_nut=cover_base) 
dat5[,LRR_nut := LRRnut_yt-LRRcont_yt]

# Selecting species again
# keep a species-site-block-trt only if has an observation in the Nut treated plots, at any point of time
dat5 <- left_join(obs.keep,dat5)

coef.npk.long <- dat5 %>% select(site_code,Taxon,block,year,year_trt,trt,num.yrs,plot_nut,cover.base_nut,cover_nut,plot_cont,cover.base_cont,cover_cont,LRRnut_yt,LRRcont_yt,LRR_nut)

rm(dat,dat2,dat3,dat4,dat5)
rm(obs.keep,cover.wide,cover,biomass_data,sites.long)

# Aggregate final products ------------------------------------------------

# Description of coef.npk.long - the base table from which metrics are aggregated
# Usual columns - site_code,Taxon,block,year,year_trt,trt
# num.yrs - number of years that this species was observed in the treated plot
# plot_nut = plot number of treated plot,
# cover.base_nut = cover of species in treated plot in base year
# cover_nut = cover of species in treated plot ,
# plot_cont = plot number of control plot,
# cover.base_cont = cover of species in control plot in base year
# cover_cont = cover of species in control plot ,
# LRRnut_yt = LRR(base=2) of cover relative to base year in treated plot,
# LRRcont_yt = LRR(base=2) of cover relative to base year in control plot,
# LRR_nut = detrended LRR(base 2) of species cover in treated vs control

#Now a slight problem with a large number of zeros in this calculated data
# dat <- coef.npk.long[cover_nut==0.01 & cover_cont==0.01 & LRR_nut!=0] %>% arrange(desc(LRR_nut))

#There are a lot of entries (14%) where the species was not seen in either control or treatment plot that year, but is only there because we added minimum cover amount to the data
# We cannot remove all of them, that would eliminate the signal of species that were present in a plot initially and then disappeared.
# But if we don't remove them, are we giving a lot of power to what could be an erroneous observation?

#Aggregate across blocks to get mean response of a species within a site
coef.npk.t <- dcast(coef.npk.long,site_code+Taxon+year+year_trt+block ~ trt,value.var='LRR_nut', fun.aggregate = mean,drop=T,fill=NA)
coef.npk.t <- rename(coef.npk.t,LRR_N = N, LRR_P = P, LRR_K = K, LRR_NPK = NPK, LRR_Fence = Fence)

coef.npk.raw <- dcast(coef.npk.long,site_code+Taxon+year+year_trt+block ~ trt,value.var='LRRnut_yt', fun.aggregate = mean,drop=T,fill=NA)
coef.npk.raw <- rename(coef.npk.raw,LRR_N = N, LRR_P = P, LRR_K = K, LRR_NPK = NPK, LRR_Fence = Fence)

#Aggregate response across years for a species
# coef.npk <- coef.npk.t%>% group_by(site_code, Taxon, block) %>% 
#   summarize(LRR_N = mean(LRR_N,na.rm=T),LRR_N_sd = sd(LRR_N,na.rm=T),nyear_N = length(unique(year_trt)),
#             LRR_P = mean(LRR_P,na.rm=T),LRR_P_sd = sd(LRR_P,na.rm=T),nyear_P = length(unique(year_trt)),
#             LRR_K = mean(LRR_K,na.rm=T),LRR_K_sd = sd(LRR_K,na.rm=T),nyear_K = length(unique(year_trt)),
#             LRR_NPK = mean(LRR_NPK,na.rm=T),LRR_NPK_sd = sd(LRR_NPK,na.rm=T),nyear_NPK = length(unique(year_trt)),
#             LRR_Fence = mean(LRR_Fence,na.rm=T),LRR_Fence_sd = sd(LRR_Fence,na.rm=T),nyear_Fence = length(unique(year_trt)))
coef.npk.t[is.na(coef.npk.t)] <- NA #replace all NaN (generated when there were no observations for a set) - with NAs

coef.npk.t <- data.table(left_join(data.frame(coef.npk.t),spsinfo))
coef.npk.raw <- data.table(left_join(data.frame(coef.npk.raw),spsinfo))

# coef.npk <- data.table(left_join(data.frame(coef.npk),spsinfo))
```

## Varying mean cover

__Parameters held constant:__

* Persistence = .2
* Initial representation in plots = 4 (all treatment plots)

__NOTE: ALL PLOTS FACETTED BY YEAR__

```{r, fig.height=15, fig.width = 10, warning = FALSE, message = FALSE}

tuning_LRRs <- function(meancov = .05,
                     persistence = .2,
                     initrep = 4){
  
  toselect = tuningtable %>% filter(
    mean_covControl >= meancov &
    (tot_obsControl / (lastyear - firstyear + 1)) >= persistence &
    pret_obsControl + pret_obsN + pret_obsP + pret_obsK >= initrep &
    pret_obsControl == 1) %>%
    select(site_code, block, Taxon, nyears)

  return(toselect)  
}


for(mc in seq(0, .5, by = .25)){
  cat(paste("Mean cover = ", mc))

  p1 = tuning_LRRs(meancov = mc, persistence = .2) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_P,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE) +
    ggtitle(paste("N - P Tradeoff, mean cover >=", mc))
  
  p2 = tuning_LRRs(meancov = mc, persistence = .2) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR K Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("N - K Tradeoff, mean cover >=", mc))
    #theme(legend.position='bottom')
  
  p3 = tuning_LRRs(meancov = mc, persistence = .2) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_P,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR K Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("P - K Tradeoff, mean cover >=", mc))
  
  grid.arrange(p1, p2, p3, nrow = 3)
}

```

## Varying persistence

__Parameters held constant:__

* Mean cover = .05
* Initial representation in plots = 4 (all treatment plots)

```{r, fig.height=15, fig.width = 10, warning = FALSE, message = FALSE}

for(pers in seq(.25, 1, by = .25)){
  
  cat(paste("Persistence = ", pers))

  p1 = tuning_LRRs(meancov = .05, persistence = pers) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_P,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE) +
    ggtitle(paste("N - P Tradeoff, persistence >=", pers))
  
  p2 = tuning_LRRs(meancov = .05, persistence = pers) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR K Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("N - K Tradeoff, persistence >=", pers))
    #theme(legend.position='bottom')
  
  p3 = tuning_LRRs(meancov = .05, persistence = pers) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_P,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR K Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("P - K Tradeoff, persistence >=", pers))
  
  grid.arrange(p1, p2, p3, nrow = 3)
}
```

## Varying initial representation

__Parameters held constant:__

* Mean cover = .05
* Persistence = .2


```{r, fig.height=15, fig.width = 10, warning = FALSE, message = FALSE}

for(initrep in seq(1, 4, by = 1)){
  
  cat(paste("Initial Representation = ", initrep))
  
  p1 = tuning_LRRs(meancov = .05, persistence = .2, initrep = initrep) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_P,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE) +
    ggtitle(paste("N - P Tradeoff, Initial representation >=", initrep))
  
  p2 = tuning_LRRs(meancov = .05, persistence = .2, initrep = initrep) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_N,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition - LRR Control',y='LRR K Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("N - K Tradeoff, Initial representation >=", initrep))
    #theme(legend.position='bottom')
  
  p3 =tuning_LRRs(meancov = .05, persistence = .2, initrep = initrep) %>% left_join(coef.npk.t) %>%
    ggplot(aes(y=LRR_P,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR K Addition - LRR Control',y='LRR P Addition - LRR Control') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("P - K Tradeoff, Initial representation >=", initrep))
  
  grid.arrange(p1, p2, p3, nrow = 3)
}
```

# Removing detrending correction:

__Parameters held constant:__

* Mean cover = .05
* Persistence = .2
* Initial representation in year 0 = 4

```{r, fig.height=15, fig.width = 10, warning = FALSE, message = FALSE}
  p1 = tuning_LRRs(meancov = .05, persistence = .2) %>% left_join(coef.npk.raw) %>%
    ggplot(aes(y=LRR_N,x=LRR_P,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition ',y='LRR P Addition') +
    guides(shape = FALSE, color = FALSE) +
    ggtitle(paste("N - P Tradeoff, No Detrending"))
  
  p2 = tuning_LRRs(meancov = .05, persistence = .2) %>% left_join(coef.npk.raw) %>%
    ggplot(aes(y=LRR_N,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR N Addition',y='LRR K Addition') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("N - K Tradeoff, No Detrending"))
    #theme(legend.position='bottom')
  
  p3 = tuning_LRRs(meancov = .05, persistence = .2) %>% left_join(coef.npk.raw) %>%
    ggplot(aes(y=LRR_P,x=LRR_K,color=functional_group,shape=functional_group,nyears)) + 
    geom_point() + geom_smooth(method='lm', se=F) +
    facet_wrap(~year_trt) +
    #geom_hline(yintercept=-1,linetype='dashed') + geom_vline(xintercept=-1,linetype='dashed') +
    #geom_hline(yintercept=1,linetype='dashed') + geom_vline(xintercept=1,linetype='dashed') +
    theme_bw() + labs(x='LRR K Addition',y='LRR P Addition') +
    guides(shape = FALSE, color = FALSE)+
    ggtitle(paste("P - K Tradeoff, No Detrending"))
  
  grid.arrange(p1, p2, p3, nrow = 3)
```

