---
title: "R Notebook"
output: html_notebook
---

A suggested way to tease apart what makes certain species more likely to respond 
positively to different nutrient enrichment treatments is through some variety 
of decision tree.

When we have a set of hierarchical predictors that range from site-level predictors
(precip, temp, species richness) to more individual-level predictors (functional 
type, relative rank abundance), decision trees can offer a reasonably intuitive 
way to model and interpet interactions between them.

# What features of a community best predict response?

- Pretrt abundance
    - Assign a zero if missing
- Functional type
- Exotic / native status
- Any sort of other functional characteristics worth adding?

```{r}
fread("../data/compmat_edited.csv")
```


# Check if previous code has any solutions

```{r}
library(tidyverse);library(data.table);library(dtplyr)

ttable <- fread("../Data/tuningtable_normalized.csv")

specscores_full <- read.csv("../Data/specscores.csv")

taxonomy <- fread('C:/Users/ebatz/Dropbox/NutNet Data/Nutnet-site-taxonomy.csv',
               stringsAsFactors = FALSE,
               na.strings = c('NA','NULL'))

site.covars <- read.csv("../Data/site_covars.csv")

site_pvals = read.csv("composition_pvalues.csv", header = TRUE, stringsAsFactors = FALSE)
head(site_pvals)
```

# Joining tables into master
```{r}
pretrt = ttable %>% group_by(site_code, Taxon, block) %>%
  summarise(pretrtcover = mean(pret_covN, pret_covControl, pret_covK, pret_covP)) %>%
  group_by(site_code, Taxon) %>%
  summarise(pretrtcover = mean(pretrtcover)) %>%
  mutate(species = gsub("\\.", " ", Taxon)) %>%
  select(-Taxon)

specscores = specscores_full %>%
  select(-X) %>%
  gather(key = "species",
         value = "coef", 
         -c(site, trt)) %>%
  mutate(species = gsub("\\.", " ", species)) %>%
  left_join(pretrt, by = c(site = "site_code", species = "species"))

taxonomy = taxonomy %>% mutate(species = gsub("\\.", " ", Taxon))

specscores = specscores %>% inner_join(taxonomy, by = c(site = "site_code", species = "species"))

specscores = specscores %>% left_join(site_pvals %>% select(site, trt_K, trt_P, trt_N))

resptable = specscores %>% left_join(site.covars, by = c(site = "site_code"))

head(resptable)
```


```{r}
df = resptable %>% filter(trt_N < .05) %>% filter(trt == "trt_N")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```
```{r}
df = resptable %>% filter(trt_P < .05) %>% filter(trt == "trt_P")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```

```{r}
df = resptable %>% filter(trt_K < .05) %>% filter(trt == "trt_K")
df$functional_group <- as.factor(df$functional_group)
df$ps_path <- as.factor(df$ps_path)
df$local_provenance <- as.factor(df$local_provenance)
df$local_lifespan <- as.factor(df$local_lifespan)
df$local_lifeform <- as.factor(df$local_lifeform)
df %>% ggplot(aes( x = log(pretrtcover),
                   y = coef,
                   color = functional_group)) + geom_point() + 
  stat_smooth(method = "lm", fullrange = TRUE, se = FALSE)
```
