---
title: "Nutrient tradeoffs among species"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(smatr)
library(tidyverse)
library(data.table); library(dtplyr)
# Run the first block of code if you want to calculate the LRRs within this script, otherwise just load the precalculated data
load('LRR_NPK_measurements_2018-08-13.RData')
```

Compiled on `r date()`

```{r Calculating LRRs, include=FALSE}
rm(list=ls())

# Read biomass data
biomass_data <- fread('C:/Users/ebatz/Dropbox/NutNet Data/comb-by-plot-09-April-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

biomass_data$site_code <- factor(biomass_data$site_code)
biomass_data$trt <- factor(biomass_data$trt)

#Choose sites with at least 3 years of data
sites <- biomass_data %>% 
  group_by(site_code,region,first_nutrient_year) %>% 
  summarise(yrs.data = length(unique(year)),fyear = min(year), lyear = max(year)) %>% 
  arrange(yrs.data)

sites.long <- as.character(sites[yrs.data >= 3,`site_code`])
rm(sites)

bmass <- biomass_data[site_code %in% sites.long]

cover <- fread('C:/Users/ebatz/Dropbox/NutNet Data/full-cover-09-April-2018.csv',
               stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))
cover.long <- cover[site_code%in%c(sites.long) & live==1,]
cover.long <- left_join(cover.long,distinct(bmass[,.(`site_code`,`region`)]))
```
```{r, include = FALSE}
#Choosing only our treatments of interest
cover.long <- cover.long[trt %in% c('Control','N','P','K','NPK','Fence')]

#Clean out unknowns, non-vascular plants
sps <- unique(cover.long$Taxon)
unwanted.sps <- sps[c(grep('UNKNOWN',sps),grep('BRYOPHYTE',sps),grep('LICHEN',sps))]
cover.long <- cover.long[!(Taxon %in% unwanted.sps | local_lifeform == 'MOSS' | functional_group == 'BRYOPHYTE')]
cover.long[functional_group=='GRAMINOID', functional_group := 'GRASS']
rm(sps,unwanted.sps)

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% 
  group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% 
  filter(yrs.data>=3)

## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% 
  select(-min.plot,-yrs.data)

cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)

# cast long into wide
cover.wide <- dcast(cover.long,site_code+year+block+plot+trt ~ Taxon,
                    value.var='max_cover', 
                    fun.aggregate = sum,drop=T,fill=0)
```


```{r, include = FALSE}
# Adding in additional info

siteinfo <- biomass_data[site_code %in% sites.long, 
                         .(`site_code`,`site_name`,`continent`,`region`,`first_nutrient_year`,`site_richness`)] %>% distinct()

spsinfo <- select(data.frame(cover.long),site_code,Family:ps_path,-live) %>% distinct()

#Summarizing some information about the species
#Max cover reached in a plot in that site, CV of cover
# coeflist.site <- left_join(coeflist.site,cover.long %>% group_by(site_code, Family, Taxon) %>% summarize(max.cover = max(max_cover), cv.cover=sd(max_cover)/mean(max_cover)))
```


```{r, include = FALSE}
# Independent measure of N,P and K response following EL 2013


# Intermediate data products
### dat2 - cover data from which LRRs are being calculated
### dat3 - baseline year data for each species-plot
### dat4 - LRR for each species plot relative to baseline
### dat5 - LRR of nut effect after detrending control

#Species selection - 	"only those species recorded at least once either pre or 
# post treatment in the nutrient treatment plot within a block"
# keep a species-site-block-trt only if has an observation in the Nut treated plots, at any point of time
obs.keep <- cover.long[trt != 'Control'] %>% group_by(site_code,Taxon,block,trt) %>% 
  summarize(num.yrs = length(unique(year))) %>% mutate(iden = paste0(site_code,Taxon,block))
# Use at the end of the procedure to only select the ones we want

dat <- cover.long[,iden := paste0(site_code,Taxon,block)]
dat <- dat[iden %in% unique(obs.keep$iden)]
obs.keep[,iden:= NULL]
#convert to wide, but only at the site-block scale, to avoid too many zeros
# Add minimum observed percent cover (0.01) to all values - to the wide table
# Convert back to long
dat2 <- data.table()
#i='cdcr.us' ; j = 1
for (i in unique(dat$site_code)) {
  subdat <- dat[site_code==i]
  for (j in unique(subdat$block)) {
    subdat2 <- subdat[block==j]
    dat2.wide <- dcast(subdat2,site_code+year+year_trt+block+plot+trt ~ Taxon,
                       value.var='max_cover', 
                       fun.aggregate = sum,
                       drop=T,
                       fill=0)
    #Now add 0.01 to all values
    dat2.wide[,7:ncol(dat2.wide)] <- dat2.wide[,7:ncol(dat2.wide)] + 0.01
    dat2.long <- melt(dat2.wide,id.vars=c(1:6),variable.name='Taxon',value.name='max_cover')
    dat2 <- bind_rows(dat2,dat2.long)
  }
}

rm(dat2.long,dat2.wide,i,j,subdat,subdat2)

# Calculate LRR of each year relative to pretreatment year ----------------

#add in cover of pretrt year for each species-plot combination
dat3 <- dat2 %>% group_by(site_code,plot,Taxon) %>% filter(year_trt == min(year_trt)) 
dat3 <- dat3[,.(`site_code`,`plot`,`Taxon`,`year_trt`,`max_cover`)] %>% rename(base_year=year_trt,cover_base=max_cover)
#dat3 <- dat3[,.(`site_code`,`plot`,`Taxon`,`base_year`=year_trt,`cover_base`=max_cover)]
#    select(site_code,plot,Taxon,`base_year`=year_trt,`cover_base`=max_cover)
# sier.us, sgs.us and hopl.us show some species as having their min(year_trt) measurement in y1
# saline.us and saana.fi are missing pretrt year in the cover dataset
#table(dat1[site_code=='hopl.us']$plot,dat1[site_code=='hopl.us']$year)
dat4 <- left_join(dat2,dat3)
dat4 <- dat4[year_trt != base_year] #remove the base year
dat4[,LRRt := log2(max_cover/cover_base)] ;
#Chose log2 in order for the coefficients to be interpretable


# Detrend from the control ------------------------------------------------

#add a column of control for each species-block in a year
dat5 <- left_join(dat4[trt!='Control'],
                  select(dat4[trt=='Control'],site_code,Taxon,year,year_trt,block,plot_cont=plot,cover_cont=max_cover,cover.base_cont=cover_base,LRRcont_yt=LRRt))
dat5 <- rename(dat5,LRRnut_yt=LRRt,plot_nut=plot,cover_nut=max_cover,cover.base_nut=cover_base) 
dat5[,LRR_nut := LRRnut_yt-LRRcont_yt]

# Selecting species again
# keep a species-site-block-trt only if has an observation in the Nut treated plots, at any point of time
dat5 <- left_join(obs.keep,dat5)

coef.npk.long <- dat5 %>% select(site_code,Taxon,block,year,year_trt,trt,num.yrs,plot_nut,cover.base_nut,cover_nut,plot_cont,cover.base_cont,cover_cont,LRRnut_yt,LRRcont_yt,LRR_nut)

rm(dat,dat2,dat3,dat4,dat5)
# rm(obs.keep,cover.wide,cover,biomass_data,sites.long)

# Aggregate final products ------------------------------------------------

# Description of coef.npk.long - the base table from which metrics are aggregated
# Usual columns - site_code,Taxon,block,year,year_trt,trt
# num.yrs - number of years that this species was observed in the treated plot
# plot_nut = plot number of treated plot,
# cover.base_nut = cover of species in treated plot in base year
# cover_nut = cover of species in treated plot ,
# plot_cont = plot number of control plot,
# cover.base_cont = cover of species in control plot in base year
# cover_cont = cover of species in control plot ,
# LRRnut_yt = LRR(base=2) of cover relative to base year in treated plot,
# LRRcont_yt = LRR(base=2) of cover relative to base year in control plot,
# LRR_nut = detrended LRR(base 2) of species cover in treated vs control

#Now a slight problem with a large number of zeros in this calculated data
#dat <- coef.npk.long[cover_nut==0.01 & cover_cont==0.01 & LRR_nut!=0] %>% arrange(desc(LRR_nut))
#hist(dat$num.yrs)
#There are a lot of entries (14%) where the species was not seen in either control or treatment plot that year, but is only there because we added minimum cover amount to the data
# We cannot remove all of them, that would eliminate the signal of species that were present in a plot initially and then disappeared.
# But if we don't remove them, are we giving a lot of power to what could be an erroneous observation?

#Adding cover + block + plot. More datapoints, but less 
coef.npk.t <- dcast(coef.npk.long,site_code+Taxon+year+year_trt+cover_cont+cover.base_cont+block ~ trt, 
                    value.var=c('LRR_nut', 'cover_nut', 'cover.base_nut'), 
                    fun.aggregate = mean,drop=T,fill=NA)

coef.npk.t <- dplyr::rename(coef.npk.t, 
                            LRR_N = LRR_nut_N, 
                            LRR_P = LRR_nut_P, 
                            LRR_K = LRR_nut_K, 
                            LRR_NPK = LRR_nut_NPK, 
                            LRR_Fence = LRR_nut_Fence)

#Aggregate response across years for a species
coef.npk <- coef.npk.t%>% group_by(site_code, Taxon) %>% 
  summarize(LRR_N = mean(LRR_N,na.rm=T),LRR_N_sd = sd(LRR_N,na.rm=T),nyear_N = length(unique(year_trt)),
            LRR_P = mean(LRR_P,na.rm=T),LRR_P_sd = sd(LRR_P,na.rm=T),nyear_P = length(unique(year_trt)),
            LRR_K = mean(LRR_K,na.rm=T),LRR_K_sd = sd(LRR_K,na.rm=T),nyear_K = length(unique(year_trt)),
            LRR_NPK = mean(LRR_NPK,na.rm=T),LRR_NPK_sd = sd(LRR_NPK,na.rm=T),nyear_NPK = length(unique(year_trt)),
            LRR_Fence = mean(LRR_Fence,na.rm=T),LRR_Fence_sd = sd(LRR_Fence,na.rm=T),nyear_Fence = length(unique(year_trt)),
            cover_cont = mean(cover_cont),
            cover.base_cont = mean(cover.base_cont))
coef.npk[is.na(coef.npk)] <- NA #replace all NaN (generated when there were no observations for a set) - with NAs

coef.npk.t <- data.table(left_join(data.frame(coef.npk.t),spsinfo))
coef.npk <- data.table(left_join(data.frame(coef.npk),spsinfo))
```

# What is the distribution of cover in the control plot?

```{r, fig.height=6,fig.width=8, echo=F}
coef.npk = na.omit(coef.npk)
b = c(-100, 0,1,5,10,20,50,200)
coef.npk$controlcover = factor(.bincode(coef.npk$cover_cont, breaks = b))
coef.npk$controlcover = plyr::revalue(coef.npk$controlcover, c("1" = "0", "2"="0-1", "3"="1-5", "4"="5-10", "5"="10-20", "6"="20-50", "7" = "50+"))

coef.npk$controlbase = factor(.bincode(coef.npk$cover.base_cont, breaks = b))
coef.npk$controlbase = plyr::revalue(coef.npk$controlbase, c("1" = "0", "2"="0-1", "3"="1-5", "4"="5-10", "5"="10-20", "6"="20-50", "7" = "50+"))

hist(coef.npk$cover_cont, main = "Histogram of Cover in Control Plot")
```


```{r}
coef.npk.t %>% ggplot(aes(x = LRR_N,
                          y = LRR_P)) +
  geom_point() +
  stat_smooth(method = "lm") +
  facet_wrap(~year)
```

# Tradeoffs facetted by cover in control plot

The following figures showed LRR values facetted by the recorded cover in control plots for a given species-plot-year combination (cover_cont). To make sure nothing was really messed up by 

Consistent with the above histogram, a large number of observations fall within the first grouping of cover values between 0 - 1%. Given our correction made earlier (+.01), I suspect that many of these correspond to zero values. As we focus more on more abundant species, the patterns seem to become less linear.

Some of these values are quite surprising. While it's reasonable to note that very rare species have LRRs of >10, LRR scores of 5 or more in very common species seem odd. With log2() transformation, an LRR of 5 corresponds to a 32-fold change.

```{r}
ggplot(coef.npk, 
       aes(y=LRR_N,
           x=LRR_P,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~controlcover) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR P Addition - LRR Control',y='LRR N Addition - LRR Control') +
  ggtitle("N-P Tradeoff Facetted by Cover in Control")

ggplot(coef.npk, 
       aes(y=LRR_N,
           x=LRR_K,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~controlcover) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR K Addition - LRR Control',y='LRR N Addition - LRR Control') +
  ggtitle("N-K Tradeoff Facetted by Cover in Control")

ggplot(coef.npk, 
       aes(y=LRR_K,
           x=LRR_P,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~controlcover) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR P Addition - LRR Control',y='LRR K Addition - LRR Control') +
  ggtitle("K-P Tradeoff Facetted by Cover in Control")
```

# Facetted by years of treatment

Strangely, I don't see that treatment effect increases that much over time. Perhaps this is just a function of the sites that have received long-term treatments?

```{r}
ggplot(coef.npk, 
       aes(y=LRR_N,
           x=LRR_P,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~nyear_N) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR P Addition - LRR Control',y='LRR N Addition - LRR Control') +
  ggtitle("N-P Tradeoff Facetted by Years of Treatment")

ggplot(coef.npk, 
       aes(y=LRR_N,
           x=LRR_K,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~nyear_N) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR K Addition - LRR Control',y='LRR N Addition - LRR Control') +
  ggtitle("N-K Tradeoff Facetted by Years of Treatment")

ggplot(coef.npk, 
       aes(y=LRR_K,
           x=LRR_P,
           color=functional_group,
           fill=functional_group)) + 
  facet_wrap(~nyear_N) + 
  geom_point(aes(color = NULL), shape = 21, alpha = .25) + 
  geom_smooth(method='lm', se=F) +
  geom_hline(yintercept=-1,linetype='dashed') + 
  geom_vline(xintercept=-1,linetype='dashed') +
  geom_hline(yintercept=1,linetype='dashed') +
  geom_vline(xintercept=1,linetype='dashed') +
  theme_bw() + 
  labs(x='LRR P Addition - LRR Control',y='LRR K Addition - LRR Control') +
  ggtitle("K-P Tradeoff Facetted by Years of Treatment")
```
