---
title: "Species Frequencies - NutNet"
output: html_notebook
---

```{r, include=FALSE}
library(smatr)
library(tidyverse)
library(data.table); library(dtplyr)

# Read biomass data
biomass_data <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/comb-by-plot-31-August-2018.csv',
                  stringsAsFactors = FALSE,
                  na.strings = c('NA','NULL'))

biomass_data$site_code <- factor(biomass_data$site_code)
biomass_data$trt <- factor(biomass_data$trt)

#Choose sites with at least 3 years of data
sites <- biomass_data %>% 
  group_by(site_code,region,first_nutrient_year) %>% 
  summarise(yrs.data = length(unique(year)),fyear = min(year), lyear = max(year)) %>% 
  arrange(yrs.data)

sites.long <- as.character(sites[yrs.data >= 3,`site_code`])
rm(sites)

bmass <- biomass_data[site_code %in% sites.long]

cover <- fread('C:/Users/ebatzer/Dropbox/NutNet Data/full-cover-31-August-2018.csv',
               stringsAsFactors = FALSE,
               na.strings = c('NA','NULL'))
cover.long <- cover[site_code%in%c(sites.long) & live==1,]
cover.long <- left_join(cover.long,distinct(bmass[,.(`site_code`,`region`)]))
cover.long <- cover.long[ , max_cover := as.numeric(max_cover)]

#Choose plots with at least 3 years of data
## This takes care of a few sites that have strange plot configurations,
## like sgs.us
site.plots.keep <- cover.long %>% 
  group_by(site_code,block,plot,trt) %>% 
  summarise(yrs.data = length(unique(year))) %>% 
  filter(yrs.data>=3)

## Some sites have multiple control plots in a block.
## Choosing only the lowest numbered plot belong to a treatment in each block.
site.plots.keep <- left_join(site.plots.keep,
                             site.plots.keep %>% group_by(site_code,block,trt) %>% 
                               summarize(min.plot = min(plot))) %>% 
  filter(plot==min.plot) %>% 
  select(-min.plot,-yrs.data)

cover.long <- left_join(site.plots.keep,cover.long)
cover.long <- droplevels(cover.long) #cover.long$site_code <- factor(cover.long$site_code)
rm(site.plots.keep)


head(cover.long)
```

# Summarising species frequencies and relative abundances

```{r, include = FALSE}
specobs = data.frame(cover.long) %>% 
  filter(trt %in% c("Control", "N", "P", "K")) %>%
  group_by(site_code, block, Taxon) %>%
  summarise(totalobs = n(),
            c_obs = sum(trt == "Control"),
            n_obs = sum(trt == "N"),
            k_obs = sum(trt == "K"),
            p_obs = sum(trt == "P"),
            meancov = mean(max_cover),
            mc_control = mean(max_cover[trt == "Control"]),
            mc_N = mean(max_cover[trt == "N"]),
            mc_K = mean(max_cover[trt == "K"]),
            mc_P = mean(max_cover[trt == "P"]))

trtcov <- data.frame(cover.long) %>% 
  filter(trt %in% c("Control", "N", "P", "K")) %>% 
  filter(year_trt == 0) %>% ungroup %>% select(-subplot) %>% 
  select(-plot) %>% 
  spread(key = trt, value = max_cover, fill=0) %>% 
  mutate(nonzero = as.numeric(Control > 0) + as.numeric(N > 0) + as.numeric(P > 0) + as.numeric(K > 0))

specobs = left_join(specobs, trtcov %>% 
                      select("site_code", "block", "Taxon", "Control", "N", "P", "K", "nonzero") %>% 
                      rename("fy_control" = "Control",
                             "fy_N" = "N",
                             "fy_P" = "P",
                             "fy_K" = "K")) %>% 
  replace(., is.na(.), 0) %>%
  left_join(cover.long %>% 
              group_by(site_code,block) %>% 
              summarise(yrs.data = length(unique(year))))
  

head(specobs)
```

# What fraction of observations do species occupy?

Following graph shows the average cover of a species with proportional representation in the total number of observations in each block over time. 

As we might expect, a very large number of species are not consistently represented in each site. Curiously, these aren't all rare species -- there are a number with relatively high mean abundances that are still only present in a very small fraction of observations. Generally, a species needs to have a mean %cover of roughly 10 percent to be represented in all observations.

```{r}
p = specobs %>% ggplot(aes(x = log10(meancov),
                       y = totalobs / (yrs.data * 4),
                       color = site_code)) +
  geom_point(alpha = .1) +
  guides(color = "none") +
  stat_smooth(se = FALSE,
              aes(color = NULL)) +
  xlab("Log10 Cover") +
  ylab("Proportion of Plots with Observations") +
  ggtitle("Plot Occupancy by Cover, Colored by Site")

for(i in seq(.1,.9, by = .1)){
    p = p + geom_hline(yintercept = i, color = "black") +
      geom_text(x = -1.2, y = i + .03, 
                label = paste(round(nrow(specobs %>% filter(totalobs / (yrs.data * 4) >= i)) / nrow(specobs), 2),
                              "data contained at cutoff", i),
                color = "black")
  }
p
```

# Occupancy at Year 0

It may be problematic if many species are not uniformly represented in all plots in year 0 if one is using the formula $Effect= log(\bar{x}_{i} / \bar{x}_{0})$, as calculated effects can be highly sensitive to whatever coefficient is used to render the LRR calculatable in situations where $\bar{x}_{0}$ is 0.

Here, we see that over 75% of the species found in our dataset fail to be represented in the first year of sampling - many calculations, then, will be largely dependent on what we choose as our correction constant. In Eric's paper, this was 0.1 - the minimum value found in some datasets, but a minimum cover of 1 might be more reasonable. "1" is the correction constant used in the recent legumes paper. 

```{r}
p = specobs %>% ggplot(aes(x = log10(meancov),
                       y = nonzero,
                       color = site_code)) +
  geom_point(alpha = .1) +
  guides(color = "none") +
  stat_smooth(se = FALSE,
              aes(color = NULL)) +
  xlab("Log10 Cover") +
  ylab("Plots (of 4 treatments) with Observations at Year Zero ") +
  ggtitle("Plot Occupancy at Year 0, Colored by Site")

for(i in seq(1, 4, by = 1)){
    p = p + geom_hline(yintercept = i, color = "black") +
      geom_text(x = -1.2, y = i + .1, 
                label = paste(round(nrow(specobs %>% filter(nonzero >= i)) / nrow(specobs), 2),
                              "data contained at cutoff", i),
                color = "black")
  }

p
```
