---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Individual Responses

To better understand why certain sites express correlations in their N, P, and K
response trajectories, we can look at the individual responses of species that
constitute these community-level effects.

From existing work on nutrient enrichment (both in NutNet and other meta-analysis),
there are a number of patterns that might be expected:

* Dominant species respond more positively to nutrient enrichment than non-dominants
* Certain functional groups (e.g. legumes) are likely to respond to P, K enrichment over others

These relationships may also give us more clues on what sort of site-level covariates
might be interested.

Here, I've taken the individual coefficients estimated from part 2 and plotted them
vs. a series of potential individual-level covariates. I've highlighted relationships
between dominance and functional group here, but a bunch more exploration is possible.

In the process of calculating these, I've produced a "tuning table" that has a number
of features that can also be used to subset the dataset prior to conducting these analyses:

* Pretrt presence / absence and cover
* Total observations over the course of sampling
* Pretrt dominance and control dominance (abundance / total observed cover)

```{r, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse); library(data.table);library(dtplyr);
library(gridExtra); library(grid); library(smatr)
library(plot3D); library(ggfortify); library(rgl)

specscores_full <- read.csv("../../Data/tradeoffs_specscores.csv", row.names = 1)

specscores_long <- specscores_full %>%
  gather(key = "Taxon", value = "Coef", - site, - trt) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon)) %>%
  mutate(Taxon = tolower(str_trim(Taxon))) %>%
  rename("site_code" = "site") 

cover <- fread('../../Data/full-cover-02-August-2019.csv',
               stringsAsFactors = FALSE,
               na.strings = c('NA','NULL'))

specfilters <- read.csv("../../Data/tradeoffs_species_filtering.csv", row.names = 1)

taxonomy <- cover %>% 
  # filter(Family != "NULL" & N_fixer != "NULL") %>%
  group_by(site_code, Taxon) %>%
  summarise(ps_path = unique(ps_path),
            functional_group = unique(functional_group)[1],
            N_fixer = unique(N_fixer)[1]) %>%
  mutate(functional_group = case_when(functional_group == "GRAMINOID" ~ "Graminoid",
                                      functional_group == "GRASS" ~ "Graminoid",
                                      functional_group == "LEGUME" ~ "Legume",
                                      functional_group == "WOODY" ~ "Woody",
                                      functional_group == "FORB" ~ "Forb")) %>%
  mutate(Taxon = gsub("\\.", " ", Taxon)) %>%
  mutate(Taxon = tolower(str_trim(Taxon)))

write.csv(taxonomy,
          "../../Data/tradeoffs_taxonomy.csv")

site_pvals = read.csv("../../Data/tradeoffs_RRPP_Pvals.csv", header = TRUE, stringsAsFactors = FALSE)
site_pvals = site_pvals %>% rename(site.K = trt_K_num, site.N = trt_N_num, site.P = trt_P_num)

dot_full = read.csv("../../Data/dot_full.csv")
```

```{r}

plot(density(specfilters$Occ_Total))

sum(specfilters$Occ_Total > 0.2) / nrow(specfilters)

full_responses = specscores_long %>% 
  inner_join(taxonomy) %>% 
  inner_join(specfilters, by = c("site_code", "Taxon"))

rm(taxonomy, cover)
```

# Fitting with SMAs

```{r}
smadat <- full_responses %>%
  ungroup() %>%
  spread(key = "trt", value = "Coef") %>%
  filter(Occ_Total > 0.33) %>%
  mutate(trt_K = scale(trt_K, scale = TRUE),
         trt_N = scale(trt_N, scale = TRUE),
         trt_P = scale(trt_P, scale = TRUE))
```


```{r}
limit = 4 

# N-P tradeoff
smafit_NP <- sma(trt_P ~ trt_N * functional_group, smadat)
NP_plot <- smafit_NP$groupsummary %>%
  ggplot() +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group),
             data = smafit_NP$groupsummary,
             size = 1) +
  geom_point(aes(x = trt_N,
                 y = trt_P,
                 color = functional_group),
             data = smadat,
             alpha = .4,
             size = 1.5) +
  xlim(-limit, limit) +
  ylim(-limit, limit) +
  xlab("Standardized N Response") +
  ylab("Standardized P Response") +
  labs(color = "Functional Group") +
  ggtitle("N-P Response Pairs") +
  theme_bw() + 
  theme(legend.position="bottom")+
  theme(plot.title = element_text(size = 12, face = "bold"))

# N-K tradeoff
smafit_NK <- sma(trt_K ~ trt_N * functional_group, smadat)
NK_plot <- smafit_NK$groupsummary %>%
  ggplot() +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group),
             data = smafit_NK$groupsummary,
             size = 1) +
  geom_point(aes(x = trt_N,
                 y = trt_K,
                 color = functional_group),
             data = smadat,
             alpha = .4,
             size = 1.5) +
  xlim(-limit, limit) +
  ylim(-limit, limit) +
  xlab("Standardized N Response") +
  ylab("Standardized K Response") +
  labs(color = "Functional Group") +
  ggtitle("N-K Response Pairs") +
  theme_bw() + 
  theme(legend.position="bottom")+
  theme(plot.title = element_text(size = 12, face = "bold"))

# P-K tradeoff
smafit_PK <- sma(trt_K ~ trt_P * functional_group, smadat)
PK_plot <- smafit_PK$groupsummary %>%
  ggplot() +
  geom_abline(aes(slope = Slope,
             intercept = Int,
             color = group),
             data = smafit_PK$groupsummary,
             size = 1) +
  geom_point(aes(x = trt_P,
                 y = trt_K,
                 color = functional_group),
             data = smadat,
             alpha = .4,
             size = 1.5) +
  xlim(-limit, limit) +
  ylim(-limit, limit) +
  xlab("Standardized P Response") +
  ylab("Standardized K Response") +
  labs(color = "Functional Group") +
  ggtitle("P-K Response Pairs")+
  theme_bw() + 
  theme(legend.position="bottom")+
  theme(plot.title = element_text(size = 12, face = "bold"))


g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(NP_plot)

p3 <- arrangeGrob(arrangeGrob(NP_plot + theme(legend.position="none"),
                         NK_plot + theme(legend.position="none"),
                         PK_plot + theme(legend.position="none"),
                         nrow=1),
             mylegend, nrow=2,heights=c(10, 1))

ggsave("../../Figures/SMAfits.pdf",
       p3,
       height = 4,
       width = 8)
```

```{r}
smafit_NK <- sma(trt_K ~ trt_N * functional_group, smadat, multcomp = TRUE, multcompmethod = "adjusted")
smafit_NP <- sma(trt_P ~ trt_N * functional_group, smadat, multcomp = TRUE, multcompmethod = "adjusted")
smafit_PK <- sma(trt_K ~ trt_P * functional_group, smadat, multcomp = TRUE, multcompmethod = "adjusted")

SMA_summary <- bind_rows(list(data.frame(pair = "NP", smafit_NP$groupsummary),
               data.frame(pair = "NK", smafit_NK$groupsummary),
               data.frame(pair = "PK", smafit_PK$groupsummary))) %>%
  select(pair, group, n, r2, pval, Slope)
  
write.csv(x = SMA_summary,
           "../../Data/SMA_summary.csv")
```


# Conducting PCA

```{r}
pc <- prcomp(scale(smadat %>% select(trt_N, trt_K, trt_P)))

cat(paste("Number of observations =", nrow(smadat)))

summary(pc)
pc$rotation

data.frame(x = c(1, 2, 3), y = summary(pc)$importance[2,]) %>%
  ggplot(aes(x = x,
             y = y)) +
  geom_point(size = 2) +
  geom_line() +
  ylim(0, 1) +
  theme_bw() +
  ylab("Proportion of Variance Explained") +
  xlab("") +
  scale_x_continuous(breaks = c(1,2,3),
                     labels = c("PC1", "PC2", "PC3"))

ggsave("../../Figures/tradeoffs_pcplot.jpeg", height = 3, width = 2)

covmat <- cor(smadat %>% select(trt_N, trt_K, trt_P))

plot3d(smadat %>% select(trt_N, trt_K, trt_P), col="black", box = TRUE, size = 8, axes = FALSE, 
       xlab = "", ylab = "", zlab = "", xlim = c(-5, 5), ylim = c(-5, 5), zlim = c(-5, 5))
ellips <- ellipse3d(covmat, 
            centre=c(0,0,0), level = 0.95)
plot3d(ellips, col = "grey", alpha = 0.2, add = TRUE, box = FALSE, axes = TRUE)
box3d()
axes3d(edges = c("x--", "y--", "z-+"))
title3d(xlab = "N Response", ylab = "P Response")
mtext3d("K Response", "z-+", line = 2)

autoplot(pc, x = 2, y = 3, data = smadat, colour = "functional_group", 
         loadings = TRUE, loadings.label = TRUE, frame = TRUE) +
  ggtitle("PC2 vs. PC3 Axis Scores")

bind_cols(data.frame(pc$x), functional_group = smadat$functional_group) %>%
  ggplot(aes(x = PC2, fill= functional_group)) +
  geom_density(alpha = .2)
```

# Within-site PCA

```{r}
site_names <- unique(smadat$site_code)

global_eigenvector <- pc$rotation[,1]

output <- list()

counter <- 1

for(site_subset in site_names){
  
  pc_dat <- smadat %>% 
              filter(site_code == site_subset) %>%
              select(trt_N, trt_K, trt_P)
  
  site_pca <- prcomp(pc_dat)
  
  prop_var <- (site_pca$sdev[1] ^ 2) /  sum(site_pca$sdev ^ 2)
  
  site_eigenvector <- site_pca$rotation[,1]
  
  site_deviance <-  acos( sum(site_eigenvector*global_eigenvector) / 
                            ( sqrt(sum(site_eigenvector * site_eigenvector)) * 
                                sqrt(sum(global_eigenvector * global_eigenvector)) ) ) * 
    (180/pi)
     
     if(site_deviance > 90){
       site_deviance = 180 - site_deviance
     }
  
  output[[counter]] <- data.frame(site_code = site_subset,
                            prop_var = prop_var,
                            site_deviance = round(site_deviance,2))
  
  counter <- counter + 1
}



p1 <- bind_rows(output) %>%
  left_join(site_pvals, by = c("site_code" = "site")) %>%
  ggplot(aes(x = prop_var)) +
  geom_histogram(color = "black") +
  theme_bw() +
  xlim(0,1) +
  xlab("Proportion of Variance") + 
  ylab("Frequency") +
  ggtitle("Variance Captured by PC1")+
  theme(plot.title = element_text(size = 12, face = "bold"))

p2 <- bind_rows(output) %>%
  left_join(site_pvals, by = c("site_code" = "site")) %>%
  ggplot(aes(x = site_deviance)) +
  geom_histogram(color = "black") +
  theme_bw() +
  xlim(0,90) +
  xlab("Angle (Deg)") + 
  ylab("Frequency") +
  ggtitle("Angle Relative to Global PC1") +
  theme(plot.title = element_text(size = 12, face = "bold"))

p3 <- arrangeGrob(arrangeGrob(p1, p2,
                         nrow=1), nrow=2,heights=c(10, 1))

ggsave("../../Figures/site_char.pdf",
       p3,
       height = 4,
       width = 6)


bind_rows(output) %>%
  left_join(site_pvals, by = c("site_code" = "site")) %>%
  ggplot(aes(x = site_deviance, y= prop_var)) +
  geom_point(color = "black") +
  theme_bw() +
  xlim(0,90) +
  xlab("Angle (Deg)") + 
  ylab("Frequency") +
  ggtitle("Angle Relative to Global PC1") +
  theme(plot.title = element_text(size = 12, face = "bold"))

```

```{r}
rep_specs <- smadat %>% group_by(Taxon) %>% summarise(count = n()) %>% filter(count > 4)

smadat %>% inner_join(rep_specs) %>%
  select(Taxon, site_code, trt_K, trt_N, trt_P) %>%
  pivot_longer(cols = c("trt_K", "trt_N", "trt_P")) %>%
  ggplot(aes(x = Taxon,
             y = value,
             color = name)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))
```

