---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse); library(vegan)
write_datsets = FALSE
```

# Deriving site-level abiotic and commmunity variables

```{r}
sitechars <- read.csv("../../Data/comb-by-plot-clim-soil-diversity-03-Jun-2020.csv",
                        stringsAsFactors = FALSE)

cover.long <- read.csv("../../Data/cover_long_tradeoffs.csv",
                       row.names = 1,
                       stringsAsFactors = FALSE)

projdat <- read.csv("../../Data/tradeoffs_dimstats.csv",
                    row.names = 1,
                    stringsAsFactors = FALSE)

taxonomy <- read.csv("../../Data/tradeoffs_taxonomy.csv", row.names = 1)

```

```{r}
calc_div = function(x){
  alpha = vegan::specnumber(x)
  gamma = vegan::specnumber(colSums(x))
  beta = gamma / alpha
  
  output = data.frame(alpha = alpha,
                      gamma = rep(gamma, length(alpha)),
                      beta = beta)
  
  return(output)
  
}
```

```{r, warning=FALSE}
site_abiotic <- sitechars %>% 
  group_by(site_code) %>%
  select(managed, burned, grazed, anthropogenic, 
         MAP_v2, MAP_WARM_Q_v2, 
         MAT_v2, TEMP_WET_Q_v2,
         RAIN_PET, PET,
         pct_C, pct_N, ppm_P, ppm_K,
         proportion_par,
         site_richness, site_introduced_richness,
         live_mass, litter_mass) %>%
  mutate_all(as.numeric) %>%
  summarise_all(function(x) mean(na.omit(x)))
```

```{r}
cover.spat <- cover.long %>% 
  filter(year_trt == 0) %>%
  select(site_code, block, plot, Taxon, max_cover) %>%
  mutate(max_cover = as.numeric(max_cover)) %>%
  group_by(site_code, block, plot, Taxon) %>%
  summarise(max_cover = sum(max_cover)) %>%
  pivot_wider(values_from = max_cover,
              names_from = Taxon,
              values_fill = list(max_cover = "0"))

sevi_subset <- cover.spat %>% filter(site_code == "sevi.us") %>%
  ungroup() %>%
  mutate(block = ceiling(plot / 8))

cover.spat[cover.spat$site_code == "sevi.us",] <- sevi_subset

spat_div = cover.spat %>% 
  group_by(site_code, block) %>%
  do(cbind(plot = .$plot, 
           calc_div(.[,-c(1:3)]))) %>%
  ungroup() %>%
  select(-plot) %>%
  group_by(site_code) %>%
  summarise_all(mean) %>%
  rename(c('alpha_spat' = 'alpha',
              'beta_spat' = 'beta',
              'gamma_spat' = 'gamma'))

```

```{r}
cover.temp <- cover.long %>% 
  filter(year_trt < 5 & 
           trt == "Control") %>% 
  select(site_code, year_trt, plot, Taxon, max_cover) %>%
  mutate(max_cover = as.numeric(max_cover)) %>%
  group_by(site_code, year_trt, plot, Taxon) %>%
  summarise(max_cover = sum(max_cover)) %>%
  pivot_wider(values_from = max_cover,
              names_from = Taxon,
              values_fill = list(max_cover = "0"))

temp_div <- cover.temp %>%
  group_by(site_code, plot) %>%
  do(cbind(plot = .$plot, 
           calc_div(.[,-c(1:3)]))) %>%
  ungroup() %>%
  select(-plot) %>%
  group_by(site_code) %>%
  summarise_all(mean) %>%
  rename(c('alpha_temp' = 'alpha',
              'beta_temp' = 'beta',
              'gamma_temp' = 'gamma'))

```

```{r}
comm_att <- cover.long %>% 
  left_join(taxonomy, by = c("site_code", "Taxon")) %>%
  filter(year_trt == 0) %>%
  group_by(site_code, block) %>%
  summarise(total_cover = sum(max_cover),
            ann_cover = sum(max_cover[lifespan == 'ANNUAL']),
            legume_cover = sum(max_cover[functional_group == 'LEGUME'])) %>%
  mutate(ann_frac = ann_cover / total_cover,
         legume_frac = legume_cover / total_cover) %>%
  ungroup() %>%
  group_by(site_code) %>%
  select(-block) %>%
  summarise_all(mean)
```

```{r}
site_table <- projdat %>% 
  select(site_code, percent_captured) %>%
  left_join(site_abiotic, by = "site_code") %>%
  left_join(comm_att, by = "site_code") %>%
  left_join(spat_div, by = "site_code") %>%
  left_join(temp_div, by = "site_code")
```

```{r}
site_table %>% 
  ggplot(aes(x = beta_temp,
             y = percent_captured)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, alpha = .5)+
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  ggrepel::geom_label_repel(aes(label = site_code))

site_table %>% 
  ggplot(aes(x = beta_spat,
             y = percent_captured)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, alpha = .5)+
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  ggrepel::geom_label_repel(aes(label = site_code))

summary(lm(percent_captured ~ beta_spat + beta_temp, data = site_table))

```

```{r}
left_join(site_table, semdat_agg, by = "site_code") %>%
  ggplot(aes(x = spatial_beta_q0,
             y = beta_spat)) +
  geom_point()+
  ggrepel::geom_label_repel(aes(label = site_code))

left_join(site_table, semdat_agg, by = "site_code") %>%
  ggplot(aes(x = temporal_beta_q0,
             y = beta_temp)) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label = site_code))
```


```{r}
anovatabs <- read.csv("../../Data/tradeoffs_RRPP_Pvals_Summary.csv",
                      row.names = 1)


site_chars <- read.csv("../../Data/tradeoffs_sitechars.csv", 
                       row.names = 1)

if(write_datsets){
  sitechars <- read.csv("../../Data/comb-by-plot-clim-soil-diversity-03-Jun-2020.csv",
                        stringsAsFactors = FALSE)
  
  sitechars_quant <- sitechars %>% 
    filter(year_trt == 0) %>%
    select(site_code, proportion_par, Ambient_PAR, Ground_PAR, 
           live_mass, unsorted_mass,
           rich, site_year_rich,
           plot_beta, inverse_simpson,
           pct_N, ppm_P, ppm_K) %>% 
    group_by(site_code) %>%
    mutate_all(as.numeric) %>%
    summarise_all(function(x) mean(na.omit(x)))
  
  sitechars_chr <- sitechars %>% 
    filter(year_trt == 0) %>%
    select(site_code, habitat, continent, country, managed, burned, grazed, anthropogenic,
           RAIN_PET, MAT_v2, MAP_v2, MAP_VAR_v2, MAT_RANGE_v2, MAP_WARM_Q_v2, TEMP_WET_Q_v2,
           elevation) %>% 
    group_by(site_code) %>%
    summarise_all(unique)
  
  spatial_het_prod <- sitechars %>%
    filter(year_trt == 0) %>%
    group_by(site_code) %>%
    summarise(spat_sd_mass = sd(na.omit(live_mass)))
  
  tradeoffs_sitechars <- left_join(sitechars_quant, sitechars_chr, by = "site_code") %>%
    left_join(spatial_het_prod, by = "site_code")

  write.csv(x = tradeoffs_sitechars, "../../Data/tradeoffs_sitechars.csv")
}

```


```{r}
full_site_chars <- anovatabs %>%
  group_by(site) %>%
  mutate(SS = SumofSquares / SumofSquares[Source == "Total"]) %>%
  select(-SumofSquares, -DF, -P_Value) %>%
  pivot_wider(names_from = Source,
              values_from = SS) %>%
  arrange(year_trt) %>%
  left_join(site_chars, by = c("site" = "site_code")) %>%
  left_join(projdat %>% select(site_code, n_taxa, percent_captured), by = c("site" = "site_code")) %>%
  arrange(percent_captured) %>% 
  mutate(fert_var = (trt_K_num + trt_P_num + trt_N_num) /
           (trt_K_num + trt_P_num + trt_N_num + Residuals)) 
```


```{r}
semdat <- read.csv("../../Data/SEM_table.csv")

semdat_agg <- semdat %>% group_by(site_code) %>% select(-block) %>% summarise_all(mean)

full_site_chars <- full_site_chars %>%
  left_join(semdat_agg, by = c("site" = "site_code")) %>%
  filter(site != 'sevi.us')

full_site_chars %>%
  ggplot(aes(x = temporal_beta_q0,
             y = percent_captured,
             color = continent)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, alpha = .5)+
  stat_smooth(method = "lm", se = FALSE, color = "black")+
  ggrepel::geom_label_repel(aes(label = site))

summary(lm(percent_captured ~ temporal_beta_q0, data = full_site_chars))
```


full_site_chars %>%
  filter(site != "sevi.us") %>%
  ggplot(aes(x = RAIN_PET,
             y = temporal_beta_q0,
             color = continent)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, color = 'black') +
  guides(color = FALSE) +
  ggrepel::geom_label_repel(aes(label = site))

full_site_chars %>%
  filter(site != "sevi.us") %>%
ggplot(aes(x = ,
             y = temporal_beta_q0,
             color = continent)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, color = 'black') +
  guides(color = FALSE) +
  ggrepel::geom_label_repel(aes(label = site))

full_site_chars %>%
  filter(site != "sevi.us") %>%
  ggplot(aes(x = sd_N / mean_N,
             y = spatial_beta_q0,
             color = continent)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, color = 'black') +
  guides(color = FALSE) +
  ggrepel::geom_label_repel(aes(label = site))



summary(lm(temporal_beta_q0 ~  MAP_WARM_Q_v2 + TEMP_WET_Q_v2, data = full_site_chars))
```


