---
title: "R Notebook"
output: html_notebook
---

```{r}
library(lavaan)
library(lavaan.survey)
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lme4) # Linear mixed effects models
library(vegan) # General ecology functions
library(data.table) # Fast reading of data tables
library(testthat) # Unit testing
library(dtplyr) # Data.table dplyr functions
library(JostDiv) # Exponentiated diversity indices (Hill Numbers)
library(gridExtra) # Multiple GGplots
library(corrplot) # Correlation plots
library(sjPlot) # Model coefficient plotting
library(tidyverse)
```

Re-run models as SEM?

Start with site-level predictors of biomass and variance in biomass

- Precipitation deficit (PET - PTo?) or function of MAP + MAT
- Soil resources (N, P, K, C, pH)
- Disturbance + management
- Invasion history?

Use these to predict site-level biomass, then also estimate variance in productivity by variance in these site-level predictors? How to best link soil resources to biomass?

- Then, link between these other drivers, biomass, and distribution of diversity within each site?

Plot partial correlations between resource heterogeneity + supply on turnover / diversity?

## Follow with plot of diversity partitioning as a function of difference in light interception

Is beta-diversity driven by changes in nestedness within each sites (species packing?) or turnover (different affinities to competitive environment / soil conditions)

## Variable definitions

### Spatial Dataset

* biomass ~ precipitation (MAT / MAT / PET - PTo) + mean soil resource supply + other factors
* spatial heterogeneity in productivity ~ some function of within-site soil resource heterogeniety + other factors
* total species pool ~ function of both biomass, heterogeneity, and other factors
* Representation in an individual plot ~ function of total species pool biomass, heterogeneity, and other factors

### Creating spatial dataset

```{r}
# Explanatory variables
explanatory_var <- read.csv("../../Data/divturnover_explanatory.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# Temporal diversity
temporal_div <- read.csv("../../Data/diversity_temporal_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# diversity
spatial_div <- read.csv("../../Data/diversity_spatial_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# General plot descriptions
plot.descriptions <- read.csv("../../Data/comb-by-plot-clim-soil-diversity-02-Aug-2019.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

```

```{r}
temporal_table <- temporal_div %>% 
  select(site_code, block, alpha_q0, beta_q0, gamma_q0) %>% 
  group_by(site_code, block) %>%
  summarise(beta_q0 = mean(beta_q0),
            alpha_q0 = mean(alpha_q0),
            gamma_q0 = mean(gamma_q0)) %>%
  rename("temporal_beta_q0" = "beta_q0",
         "temporal_alpha_q0" = "alpha_q0",
         "temporal_gamma_q0" = "gamma_q0")

spatial_table <- spatial_div %>% 
  select(site_code, block, alpha_q0, beta_q0, gamma_q0) %>% 
  group_by(site_code, block) %>%
  summarise(beta_q0 = mean(beta_q0),
            alpha_q0 = mean(alpha_q0),
            gamma_q0 = mean(gamma_q0)) %>%
  rename("spatial_beta_q0" = "beta_q0",
         "spatial_alpha_q0" = "alpha_q0",
         "spatial_gamma_q0" = "gamma_q0")

biomass_summary_spatial <- plot.descriptions %>% 
  filter(year_trt == 0) %>% 
  select(site_code, block, total_mass, proportion_par, Ambient_PAR, live_mass) %>% 
  group_by(site_code, block) %>%
  mutate(proportion_par = proportion_par * 100) %>%
  summarise(spatial_mean_biomass = mean(na.omit(total_mass)),
            spatial_cv_biomass = sd(na.omit(total_mass)) /  mean(na.omit(total_mass)),
            spatial_mean_prod = mean(na.omit(live_mass)),
            spatial_cv_prod = sd(na.omit(live_mass)) / mean(na.omit(live_mass)),
            spatial_mean_par = mean(100 - proportion_par),
            spatial_sd_par = sd(100 - proportion_par),
            spatial_ambient_par = mean(Ambient_PAR))

biomass_summary_temporal <- plot.descriptions %>% 
  filter(year_trt < 5 & trt == "Control" & !is.na(first_nutrient_year)) %>%
  select(site_code, block, total_mass, proportion_par, Ambient_PAR, live_mass) %>%
  group_by(site_code, block) %>%
  mutate(proportion_par = proportion_par * 100) %>%
  summarise(temporal_mean_biomass = mean(na.omit(total_mass)),
            temporal_cv_biomass = sd(na.omit(total_mass)) / mean(na.omit(total_mass)),
            temporal_mean_prod = mean(na.omit(live_mass)),
            temporal_cv_prod = sd(na.omit(live_mass)) / mean(na.omit(live_mass)),
            temporal_mean_par = mean(100 - proportion_par),
            temporal_sd_par = sd(100 - proportion_par),
            temporal_ambient_par = mean(Ambient_PAR)) %>%
  # group_by(site_code, block) %>%
  # summarise(temporal_mean_biomass = mean(na.omit(temporal_mean_biomass)),
  #           temporal_sd_biomass = mean(na.omit(temporal_sd_biomass)),
  #           temporal_mean_par = mean(na.omit(temporal_mean_par)),
  #           temporal_sd_par = sd(na.omit(temporal_sd_par)),
  #           temporal_ambient_par = mean(na.omit(temporal_ambient_par))) %>%
  mutate(temporal_cv_biomass = ifelse(site_code == "ethass.au", NA, temporal_cv_biomass))


explanatory_cvs <- explanatory_var %>%
  transmute(site_code = site_code,
            block = block, 
            cv_N = sd_N / mean_N,
            cv_C = sd_C / mean_C,
            cv_P = sd_P / mean_P,
            cv_K = sd_K / mean_K,
            cv_ppt = sd_ppt / mean_ppt,
            cv_pet = sd_pet / mean_pet)

SEM_table <- spatial_table %>% 
  full_join(temporal_table, by = c("site_code", "block")) %>%
  left_join(biomass_summary_spatial, by = c("site_code", "block")) %>% 
  left_join(biomass_summary_temporal, by = c("site_code", "block")) %>%
  left_join(explanatory_var, by = c("site_code", "block")) %>%
  left_join(explanatory_cvs, by = c("site_code", "block"))

write.csv(x = SEM_table,
          "../../Data/SEM_table.csv", 
          row.names = FALSE)

SEM_table %>%
  select(-anthropogenic, - burned, -grazed, -managed) %>%
  gather(key = "sd", value = "val", -site_code) %>%
  ggplot(aes(x = val)) +
  geom_histogram(bins = 30, fill = "white", color = "black") +
  facet_wrap(~sd, scales = "free")

ggsave("../../Figures/check_distributions.pdf", height = 15, width = 25)
```

```{r}
spatial_SEM_dat = SEM_table %>% select(site_code, 
                                       spatial_beta_q0, spatial_gamma_q0, 
                                       spatial_cv_biomass, spatial_cv_prod, 
                                       managed, burned, grazed,
                                       cv_N, cv_C, cv_K, cv_P, sd_pH) %>%
  na.omit(.) %>%
  filter(cv_P != 0) %>%
  mutate(managed = factor(managed, levels = c(0,1)),
         burned = factor(burned, levels = c(0,1)),
         grazed = factor(grazed, levels = c(0,1))) %>%
  mutate_if(is.numeric, log)


soil_composite_model <- lm(spatial_cv_prod ~ cv_N + cv_C + cv_K + cv_P + sd_pH, data = spatial_SEM_dat)
soil_composite_coefs <- coef(soil_composite_model)[2:5]
soil_composite <- soil_composite_coefs * spatial_SEM_dat %>% select(cv_N, cv_C, cv_K, cv_P, sd_pH)


```

# Need to check out derr.au



























```{r}
SEM_subset <- SEM_table %>% select(spatial_alpha_q0, spatial_beta_q0, spatial_gamma_q0, spatial_cv_biomass, spatial_sd_par, spatial_cv_prod) %>% ungroup()

SEM_subset <- na.omit(SEM_subset)

cor(SEM_subset %>% select(spatial_cv_biomass, spatial_sd_par, spatial_cv_prod, spatial_beta_q0, spatial_gamma_q0))

SEM_subset %>% ggplot(aes(x = log(spatial_beta_q0),
                             y = log(spatial_gamma_q0))) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) +
  xlab("Spatial Gamma Diversity (log scale)") +
  ylab("Spatial Beta Diversity (log scale)")

SEM_subset$gamma_resid <- resid(lm(log(spatial_beta_q0) ~ log(spatial_gamma_q0), data = SEM_subset))

SEM_subset %>% ggplot(aes(x = log(spatial_sd_par),
                             y = log(spatial_gamma_q0),
                          color = site_code)) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  xlab("Spatial Light Heterogeneity") +
  ylab("Spatial Gamma Diversity") +
  guides(color = FALSE)

SEM_subset %>% ggplot(aes(x = log(spatial_cv_biomass),
                             y = log(spatial_gamma_q0),
                          color = site_code)) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  xlab("Spatial Biomass Heterogeneity") +
  ylab("Spatial Gamma Diversity") +
  guides(color = FALSE)

SEM_subset %>% ggplot(aes(x = log(spatial_cv_biomass),
                             y = scale(gamma_resid),
                          color = site_code)) +
  
    geom_point() + 
    stat_smooth(method = "lm", se = FALSE) +
    stat_smooth(method = "lm", se = FALSE, color = "black") +
    xlab("Spatial Biomass Heterogeneity") +
    ylab("Spatial Beta Diversity") +
    guides(color = FALSE)

SEM_subset %>% ggplot(aes(x = log(spatial_sd_par),
                          y = log(spatial_beta_q0),
                          color = site_code)) +
    geom_point() + 
    stat_smooth(method = "lm", se = FALSE) +
    stat_smooth(method = "lm", se = FALSE, color = "black") +
    xlab("Spatial Light Heterogeneity") +
    ylab("Spatial Beta Diversity") +
    guides(color = FALSE)

SEM_subset %>% ggplot(aes(x = log(spatial_cv_biomass),
                          y = log(spatial_cv_prod),
                          color = site_code)) +
    geom_point() + 
    stat_smooth(method = "lm", se = FALSE) +
    stat_smooth(method = "lm", se = FALSE, color = "black") +
    xlab("Spatial Light Heterogeneity") +
    ylab("Spatial Beta Diversity") +
    guides(color = FALSE)
```


```{r}
SEM_table$gamma_resid <- resid(lm(log(spatial_gamma_q0) ~ log(spatial_alpha_q0), data = SEM_table))

gamma_resids <- SEM_table %>% ggplot(aes(x = temporal_cv_biomass,
                             y = scale(gamma_resid))) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) +
  xlab("Spatial Biomass Heterogeneity") +
  ylab("Spatial Beta Diversity")

log_beta <- SEM_table %>% ggplot(aes(x = temporal_cv_biomass,
                             y = scale(log(spatial_beta_q0)))) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) +
  xlab("Spatial Biomass Heterogeneity") +
  ylab("Spatial Beta Diversity")
grid.arrange(gamma_resids, log_beta, nrow = 1)
```

```{r}
# MAP
hist(MAP_v2) # log transform

# MAT
hist(MAT_v2) # No transformation

# ann_frac
hist(log(ann_frac + 1)) # Log transform? This is a tricky one, because there are lots of zeroes. Maybe don't use

# Soil variables (already log-transformed)
hist(mean_N)

# Spatial biomass
hist(spatial_mean_biomass) # Definitely log transform

# Temporal biomass
hist(temporal_mean_biomass) # Definitely log transform
```


```{r}
# Create data object and rename variables for clarity
site.sem.dat <- data.frame()
site.sem.dat<- spatial_table$SiteCode

# Richness
site.sem.dat$SiteRich <- log(spatial_table$spatial_gamma_q0)
site.sem.dat$SiteBeta <- log(spatial_table$spatial_beta_q0)
  
# Soil resources (means are log-transformed)
site.sem.dat$mean_N <- spatial_table$mean_N
site.sem.dat$mean_C <- spatial_table$mean_C
site.sem.dat$mean_P <- spatial_table$mean_P
site.sem.dat$mean_K <- spatial_table$mean_K
site.sem.dat$mean_ph <- spatial_table$mean_pH

# Heterogeneity in soil resources
site.sem.dat$var_N <- spatial_table$sd_N
site.sem.dat$var_C <- spatial_table$sd_C
site.sem.dat$var_P <- spatial_table$sd_P
site.sem.dat$var_K <- spatial_table$sd_K
site.sem.dat$var_ph <- spatial_table$sd_pH

# Climate
site.sem.dat$MAP  <- log(spatial_table$MAP_v2)
site.sem.dat$MAT <- spatial_table$MAT_v2

# PAR
site.sem.dat$light_sd <- spatial_table$spatial_sd_par
site.sem.dat$light_mean <- spatial_table$spatial_mean_par

# Biomass and Site Covariates
site.sem.dat$SiteBiomass <- spatial_table$spatial_mean_biomass
site.sem.dat$SpatialHet <- spatial_table$spatial_sd_biomass
site.sem.dat$AnthroDisturbance <- spatial_table$anthropogenic #anthropogenic disturbance
site.sem.dat$Grazing <- spatial_table$grazed

# Simple SEM
sitemod <- '# Latent variables
            SoilHeterogeneity =~ var_ph + var_N + var_P + var_K
          
            # Regressions
            SpatialHet ~ SoilHeterogeneity + Grazing + AnthroDisturbance
            light_sd ~ SpatialHet 
            SiteRich ~ SpatialHet + Grazing + AnthroDisturbance
            SiteBeta ~ SiteRich + light_sd'

sitemod.fit <- sem(sitemod, data=as.data.frame(site.sem.dat), fixed.x=FALSE, meanstructure = TRUE)  # run model

summary(sitemod.fit)

```















```{r}
sitemod <- '# Latent variables
            SoilResources =~ mean_N + mean_P + mean_K 
            SoilHeterogeneity =~ var_N + var_P + var_K

            # Regressions
            SiteBiomass ~ SoilResources + MAP + Grazing + AnthroDisturbance
            SpatialHet ~ SoilHeterogeneity + Grazing + AnthroDisturbance
            light_sd ~ SpatialHet + Grazing + AnthroDisturbance
            SiteRich ~ SiteBiomass + light_sd + Grazing + AnthroDisturbance
            SiteBeta ~ SiteRich + light_sd + SoilHeterogeneity

            # Residual Covariances
            SoilResources ~~ SoilHeterogeneity
            mean_N ~~ SpatialHet
            '



#exp.design_Ctrl <- svydesign(ids=~site_code, nest=TRUE, data=dat_SEM_Ctrl)
sitemod.fit <- sem(sitemod, data=as.data.frame(site.sem.dat), fixed.x=FALSE, meanstructure = TRUE)  # run model


fit_Ctrl <- sem(full.model.TTO, data = dat_SEM_Ctrl, estimator="MLM")
exp.design_Ctrl <- svydesign(ids=~site_code, nest=TRUE, data=dat_SEM_Ctrl)
survey.fit.plot_Ctrl <- lavaan.survey(lavaan.fit=fit_Ctrl, survey.design=exp.design_Ctrl)
summary(survey.fit.plot_Ctrl, fit.measures=T, standardized=T, rsquare=T)

```

```{r}
summary(sitemod.fit, fit.measures = TRUE)

standardizedSolution(sitemod.fit) %>% filter(pvalue < 0.10)
```


```{r}
semPlot::semPlotModel(sitemod.fit)
```

