---
title: "R Notebook"
output: html_notebook
---
```{r}
library(lavaan)
library(lavaan.survey)
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lme4) # Linear mixed effects models
library(vegan) # General ecology functions
library(data.table) # Fast reading of data tables
library(testthat) # Unit testing
library(dtplyr) # Data.table dplyr functions
library(JostDiv) # Exponentiated diversity indices (Hill Numbers)
library(gridExtra) # Multiple GGplots
library(corrplot) # Correlation plots
library(sjPlot) # Model coefficient plotting
library(tidyverse)
```

Re-run models as SEM?

Start with site-level predictors of biomass and variance in biomass

- Precipitation deficit (PET - PTo?) or function of MAP + MAT
- Soil resources (N, P, K, C, pH)
- Disturbance + management
- Invasion history?

Use these to predict site-level biomass, then also estimate variance in productivity by variance in these site-level predictors? How to best link soil resources to biomass?

- Then, link between these other drivers, biomass, and distribution of diversity within each site?

Plot partial correlations between resource heterogeneity + supply on turnover / diversity?

## Follow with plot of diversity partitioning as a function of difference in light interception

Is beta-diversity driven by changes in nestedness within each sites (species packing?) or turnover (different affinities to competitive environment / soil conditions)

## Variable definitions

### Spatial Dataset

* biomass ~ precipitation (MAT / MAT / PET - PTo) + mean soil resource supply + other factors
* spatial heterogeneity in productivity ~ some function of within-site soil resource heterogeniety + other factors
* total species pool ~ function of both biomass, heterogeneity, and other factors
* Representation in an individual plot ~ function of total species pool biomass, heterogeneity, and other factors

### Creating spatial dataset

```{r}
# Explanatory variables
explanatory_var <- read.csv("../../Data/divturnover_explanatory.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# Temporal diversity
temporal_div <- read.csv("../../Data/diversity_temporal_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# diversity
spatial_div <- read.csv("../../Data/diversity_spatial_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# General plot descriptions
plot.descriptions <- read.csv("../../Data/comb-by-plot-clim-soil-diversity-02-Aug-2019.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

```

```{r}
temporal_table <- temporal_div %>% 
  select(site_code, block, alpha_q0, beta_q0, gamma_q0) %>% 
  group_by(site_code, block) %>%
  summarise(beta_q0 = mean(beta_q0),
            alpha_q0 = mean(alpha_q0),
            gamma_q0 = mean(gamma_q0)) %>%
  rename("temporal_beta_q0" = "beta_q0",
         "temporal_alpha_q0" = "alpha_q0",
         "temporal_gamma_q0" = "gamma_q0")

spatial_table <- spatial_div %>% 
  select(site_code, block, alpha_q0, beta_q0, gamma_q0) %>% 
  group_by(site_code, block) %>%
  summarise(beta_q0 = mean(beta_q0),
            alpha_q0 = mean(alpha_q0),
            gamma_q0 = mean(gamma_q0)) %>%
  rename("spatial_beta_q0" = "beta_q0",
         "spatial_alpha_q0" = "alpha_q0",
         "spatial_gamma_q0" = "gamma_q0")

biomass_summary_spatial <- plot.descriptions %>% 
  filter(year_trt == 0) %>% 
  select(site_code, block, total_mass, proportion_par, Ambient_PAR, live_mass) %>% 
  group_by(site_code, block) %>%
  mutate(proportion_par = proportion_par * 100) %>%
  summarise(spatial_mean_biomass = mean(na.omit(total_mass)),
            spatial_cv_biomass = sd(na.omit(total_mass)) /  mean(na.omit(total_mass)),
            spatial_mean_prod = mean(na.omit(live_mass)),
            spatial_cv_prod = sd(na.omit(live_mass)) / mean(na.omit(live_mass)),
            spatial_mean_par = mean(100 - proportion_par),
            spatial_sd_par = sd(100 - proportion_par),
            spatial_ambient_par = mean(Ambient_PAR))

biomass_summary_temporal <- plot.descriptions %>% 
  filter(year_trt < 5 & trt == "Control" & !is.na(first_nutrient_year)) %>%
  select(site_code, block, total_mass, proportion_par, Ambient_PAR, live_mass) %>%
  group_by(site_code, block) %>%
  mutate(proportion_par = proportion_par * 100) %>%
  summarise(temporal_mean_biomass = mean(na.omit(total_mass)),
            temporal_cv_biomass = sd(na.omit(total_mass)) / mean(na.omit(total_mass)),
            temporal_mean_prod = mean(na.omit(live_mass)),
            temporal_cv_prod = sd(na.omit(live_mass)) / mean(na.omit(live_mass)),
            temporal_mean_par = mean(100 - proportion_par),
            temporal_sd_par = sd(100 - proportion_par),
            temporal_ambient_par = mean(Ambient_PAR)) %>%
  # group_by(site_code, block) %>%
  # summarise(temporal_mean_biomass = mean(na.omit(temporal_mean_biomass)),
  #           temporal_sd_biomass = mean(na.omit(temporal_sd_biomass)),
  #           temporal_mean_par = mean(na.omit(temporal_mean_par)),
  #           temporal_sd_par = sd(na.omit(temporal_sd_par)),
  #           temporal_ambient_par = mean(na.omit(temporal_ambient_par))) %>%
  mutate(temporal_cv_biomass = ifelse(site_code == "ethass.au", NA, temporal_cv_biomass))


explanatory_cvs <- explanatory_var %>%
  transmute(site_code = site_code,
            block = block, 
            cv_N = sd_N / mean_N,
            cv_C = sd_C / mean_C,
            cv_P = sd_P / mean_P,
            cv_K = sd_K / mean_K,
            cv_ppt = sd_ppt / mean_ppt,
            cv_pet = sd_pet / mean_pet)

SEM_table <- spatial_table %>% 
  full_join(temporal_table, by = c("site_code", "block")) %>%
  left_join(biomass_summary_spatial, by = c("site_code", "block")) %>% 
  left_join(biomass_summary_temporal, by = c("site_code", "block")) %>%
  left_join(explanatory_var, by = c("site_code", "block")) %>%
  left_join(explanatory_cvs, by = c("site_code", "block"))

write.csv(x = SEM_table,
          "../../Data/SEM_table.csv", 
          row.names = FALSE)

# distr_plot <- SEM_table %>%
#   select(-anthropogenic, - burned, -grazed, -managed) %>%
#   gather(key = "sd", value = "val", -site_code) %>%
#   ggplot(aes(x = val)) +
#   geom_histogram(bins = 30, fill = "white", color = "black") +
#   facet_wrap(~sd, scales = "free")
# 
# ggsave(plot = distr_plot, 
#        "../../Figures/check_distributions.pdf", height = 15, width = 25)
```

```{r}
spatial_SEM_dat = SEM_table %>% select(site_code, 
                                       spatial_beta_q0, spatial_gamma_q0, 
                                       spatial_cv_biomass, spatial_cv_prod, spatial_sd_par,
                                       managed, burned, grazed,
                                       cv_N, cv_C, cv_K, cv_P, sd_pH) %>%
  na.omit(.) %>%
  filter(cv_P != 0) %>%
  mutate(managed = factor(managed, levels = c(0,1)),
         burned = factor(burned, levels = c(0,1)),
         grazed = factor(grazed, levels = c(0,1))) %>%
  mutate_if(is.numeric, log) %>%
  ungroup()


soil_composite_model <- lm(spatial_cv_prod ~ cv_N + cv_C + cv_K + cv_P + sd_pH, data = spatial_SEM_dat)
summary(soil_composite_model)

soil_composite_coefs <- coef(soil_composite_model)[2:6]
soil_composite <- soil_composite_coefs[1] * spatial_SEM_dat$cv_N +
                          soil_composite_coefs[2] * spatial_SEM_dat$cv_C + 
                          soil_composite_coefs[3] * spatial_SEM_dat$cv_K +
                          soil_composite_coefs[4] * spatial_SEM_dat$cv_P +
                          soil_composite_coefs[5] * spatial_SEM_dat$sd_pH

plot(spatial_SEM_dat$spatial_cv_prod ~ soil_composite)

biomass_composite_model <- lm(spatial_cv_biomass ~ spatial_cv_prod + managed + grazed + burned, data = spatial_SEM_dat)
summary(biomass_composite_model)
biomass_composite_coefs <- coef(biomass_composite_model)[3:5]
biomass_composite <- biomass_composite_coefs[1] * (as.integer(spatial_SEM_dat$managed) - 1) +
  biomass_composite_coefs[2] * (as.integer(spatial_SEM_dat$grazed) - 1) +
  biomass_composite_coefs[3] * (as.integer(spatial_SEM_dat$burned) - 1)

plot(resid(lm(spatial_SEM_dat$spatial_cv_biomass ~ spatial_SEM_dat$spatial_cv_prod)) ~ biomass_composite)

spatial_SEM_dat = bind_cols(spatial_SEM_dat, soil_composite = soil_composite, biomass_composite = biomass_composite)
```

```{r}
sitemod <- '# Composite variables
spatial_cv_prod ~ soil_composite
spatial_cv_biomass ~ spatial_cv_prod + biomass_composite
spatial_beta_q0 ~ spatial_cv_biomass + spatial_gamma_q0

# Correlated error
spatial_cv_prod ~~ spatial_cv_biomass
'

sitemod.fit <- sem(sitemod, data=spatial_SEM_dat, fixed.x=FALSE, meanstructure = TRUE)  # run model
summary(sitemod.fit, fit.measures = TRUE)

standardizedSolution(sitemod.fit) %>% filter(pvalue < 0.10)
```

