---
title: "R Notebook"
output: html_notebook
---

# Initial Exploration of Resource Supply / Heterogeneity Relationships

See other markdown files for calculation of diversity statistics (DivTurnover_divstats.Rmd) and environmental characteristics (DivTurnover_Envtl.Rmd).

Resource Supply includes standardized abundance of:

* N
* P
* K 
* C

Initial exploration focuses on two key aspects of resource supply in heterogeneity, both in resource supply (soil resources) and temporal resource supply (climatic resources)

# Loading packages

```{r, message = FALSE, echo = FALSE}
library(ggplot2) # Figure plotting
library(tidyverse) # Data manipulation
library(lme4) # Linear mixed effects models
library(vegan) # General ecology functions
library(data.table) # Fast reading of data tables
library(testthat) # Unit testing
library(dtplyr) # Data.table dplyr functions
library(JostDiv) # Exponentiated diversity indices (Hill Numbers)
library(gridExtra) # Multiple GGplots
library(corrplot) # Correlation plots
library(sjPlot) # Model coefficient plotting

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8.5, fig.height = 5)
```

# Reading in Datasets

```{r}
# Explanatory variables
explanatory_var <- read.csv("../Data/divturnover_explanatory.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# Temporal diversity
temporal_div <- read.csv("../Data/diversity_temporal_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# diversity
spatial_div <- read.csv("../Data/diversity_spatial_full.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

# General plot descriptions
plot.descriptions <- read.csv("../data/comb-by-plot-clim-soil-diversity-02-Aug-2019.csv",
                           stringsAsFactors = FALSE,
                           na.strings = c('NA','NULL'))

```

# Assembling dataset for full model

```{r}
full_table <- spatial_div %>% 
  select(site_code, alpha_q0, beta_q0,gamma_q0, alpha_q2, beta_q2, gamma_q2, plot) %>% 
  group_by(site_code) %>%
  summarise(beta_q0 = mean(beta_q0),
            alpha_q0 = mean(alpha_q0),
            gamma_q0 = mean(gamma_q0),
            beta_q2 = mean(beta_q2),
            alpha_q2 = mean(alpha_q2),
            gamma_q2 = mean(gamma_q2)) %>%
  rename("spatial_beta_q0" = "beta_q0",
         "spatial_alpha_q0" = "alpha_q0",
         "spatial_gamma_q0" = "gamma_q0",
         "spatial_beta_q2" = "beta_q2",
         "spatial_alpha_q2" = "alpha_q2",
         "spatial_gamma_q2" = "gamma_q2") %>%
  full_join(temporal_div %>%
              select(site_code, alpha_q0, beta_q0,gamma_q0, alpha_q2, beta_q2, gamma_q2, plot) %>% 
              group_by(site_code) %>%
              summarise(beta_q0 = mean(beta_q0),
                        alpha_q0 = mean(alpha_q0),
                        gamma_q0 = mean(gamma_q0),
                        beta_q2 = mean(beta_q2),
                        alpha_q2 = mean(alpha_q2),
                        gamma_q2 = mean(gamma_q2)) %>%
              rename("temporal_beta_q0" = "beta_q0",
                     "temporal_alpha_q0" = "alpha_q0",
                     "temporal_gamma_q0" = "gamma_q0",
                     "temporal_beta_q2" = "beta_q2",
                     "temporal_alpha_q2" = "alpha_q2",
                     "temporal_gamma_q2" = "gamma_q2")) %>%
  left_join(explanatory_var) %>% select(-X)

biomass_summary_spatial <- plot.descriptions %>% 
  filter(year_trt == 0) %>% 
  select(site_code, total_mass) %>%
  group_by(site_code) %>%
  summarise(spatial_mean_biomass = mean(na.omit(total_mass)),
            spatial_cv_biomass = sd(na.omit(total_mass)) / mean(na.omit(total_mass)))

biomass_summary_temporal <- plot.descriptions %>% 
  filter(year_trt < 5 & trt == "Control" & !is.na(first_nutrient_year)) %>% 
  select(site_code, plot, total_mass) %>%
  group_by(site_code, plot) %>%
  summarise(temporal_mean_biomass = mean(na.omit(total_mass)),
            temporal_cv_biomass = sd(na.omit(total_mass)) / mean(na.omit(total_mass))) %>%
  filter(!is.na(temporal_cv_biomass))
  # summarise(temporal_mean_biomass = mean(temporal_mean_biomass),
  #          temporal_cv_biomass =  mean(temporal_cv_biomass))
 
full_table <- full_table %>% left_join(biomass_summary_spatial) %>% left_join(biomass_summary_temporal)
```

# Basic summary statistics:

## Spatial alpha, beta, and gamma diversity

## Q = 0

```{r}
p1 <- ggplot(aes(x = spatial_gamma_q0,
                 y = spatial_alpha_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Gamma")

p2 <- ggplot(aes(x = spatial_gamma_q0,
                 y = spatial_beta_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Gamma")

p3 <- ggplot(aes(x = spatial_beta_q0,
                 y = spatial_alpha_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Alpha ~ Beta")

grid.arrange(p1, p2, p3, nrow = 1)
```

## Q = 2

```{r}
p1 <- ggplot(aes(x = spatial_gamma_q2,
                 y = spatial_alpha_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Gamma")

p2 <- ggplot(aes(x = spatial_gamma_q2,
                 y = spatial_beta_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Gamma")

p3 <- ggplot(aes(x = spatial_beta_q2,
                 y = spatial_alpha_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Alpha ~ Beta")

grid.arrange(p1, p2, p3, nrow = 1)
```

# Temporal alpha, beta, and gamma diversity

## Q = 0

```{r}
p1 <- ggplot(aes(x = temporal_gamma_q0,
                 y = temporal_alpha_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Gamma")

p2 <- ggplot(aes(x = temporal_gamma_q0,
                 y = temporal_beta_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Gamma")

p3 <- ggplot(aes(x = temporal_beta_q0,
                 y = temporal_alpha_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Alpha ~ Beta")

grid.arrange(p1, p2, p3, nrow = 1)
```

## Q = 2

```{r}
p1 <- ggplot(aes(x = temporal_gamma_q2,
                 y = temporal_alpha_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Gamma")

p2 <- ggplot(aes(x = temporal_gamma_q2,
                 y = temporal_beta_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Gamma")

p3 <- ggplot(aes(x = temporal_beta_q2,
                 y = temporal_alpha_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Alpha ~ Beta")

grid.arrange(p1, p2, p3, nrow = 1)
```


# Correlations between temporal and spatial diversity

## Q = 0

```{r}
p1 <- ggplot(aes(x = spatial_alpha_q0,
                 y = temporal_alpha_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Alpha")

p2 <- ggplot(aes(x = spatial_beta_q0,
                 y = temporal_beta_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Beta")

p3 <- ggplot(aes(x = spatial_gamma_q0,
                 y = temporal_gamma_q0,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Gamma ~ Gamma")

grid.arrange(p1, p2, p3, nrow = 1)
```

## Q = 2

```{r}
p1 <- ggplot(aes(x = spatial_alpha_q2,
                 y = temporal_alpha_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE) +
  ggtitle("Alpha ~ Gamma")

p2 <- ggplot(aes(x = spatial_beta_q2,
                 y = temporal_beta_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Beta ~ Gamma")

p3 <- ggplot(aes(x = spatial_gamma_q2,
                 y = temporal_gamma_q2,
                 color = site_code),
             data = full_table) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, aes(color = NULL)) +
  guides(color = FALSE)+
  ggtitle("Alpha ~ Beta")

grid.arrange(p1, p2, p3, nrow = 1)
```

# Model Fitting

```{r}
mod_beta_spatial = lm(scale(spatial_beta_q0) ~ scale(spatial_gamma_q0) + scale(spatial_mean_biomass) + scale(spatial_cv_biomass) , data = full_table)
# summary(mod_beta_spatial)
```

```{r}
mod_alpha_spatial = lm(scale(spatial_alpha_q0) ~ scale(spatial_gamma_q0) + scale(spatial_mean_biomass) + scale(spatial_cv_biomass), data = full_table)
# summary(mod_alpha_spatial)
```

```{r}
mod_beta_temporal = lm(scale(temporal_beta_q0) ~ scale(temporal_gamma_q0) + scale(temporal_mean_biomass) + scale(temporal_cv_biomass) , data = full_table)
# summary(mod_beta_temporal)
```

```{r}
mod_alpha_temporal = lm(scale(temporal_alpha_q0) ~ scale(temporal_gamma_q0) + scale(temporal_mean_biomass) + scale(temporal_cv_biomass), data = full_table)
# summary(mod_alpha_temporal)
```

# Plotting Model Coefficients:

```{r}
# draw a dot-and-whisker plot
p1 = plot_model(mod_alpha_spatial) 
p2 = plot_model(mod_alpha_temporal) 
p3 = plot_model(mod_beta_spatial)
p4 = plot_model(mod_beta_temporal)

grid.arrange(p1,p2,p3,p4, nrow = 1)

bind_rows(p1$data,
          p2$data,
          p3$data,
          p4$data) %>%
  mutate(lev = c(rep("Spatial Plot Richness", nrow(p1$data)), 
               rep("Temporal Plot Richness", nrow(p2$data)), 
               rep("Spatial Turnover", nrow(p3$data)), 
               rep("Temporal Turnover", nrow(p4$data))),
         alpha_val = as.numeric(if_else(p.value < 0.05, 1, .3))) %>%
  mutate(term = case_when(grepl("gamma", term) ~ "Total Species Pool",
            grepl("mean", term) ~ "Total Biomass",
            grepl("cv", term) ~ "CV of Biomass")) %>%
  ggplot(aes(x = estimate,
             y = term,
             color = estimate > 0,
             fill = estimate > 0,
             alpha = I(alpha_val))) + 
  geom_vline(xintercept = 0) +
  geom_segment(aes(x = conf.low,
                 xend = conf.high,
                 y = term,
                 yend = term),
               size = 1) +
  geom_point(size = 5, pch = 21, color = "black") +
  facet_wrap(~lev, nrow = 2) +
  theme_bw() +
  xlab("Standardized Coefficient Estimate") +
  ylab("Coefficient Name") +
  guides(color = FALSE,
         fill = FALSE) +
  ggtitle("Model Coefficient Effects")
```

# Predicting biomass

```{r}
mean_site_biomass <- plot.descriptions

biomass_model <-  plot.descriptions %>% 
  filter(year_trt == 0) %>% 
  group_by(site_code) %>%
  summarise(spatial_mean_biomass = mean(na.omit(total_mass)),
            spatial_cv_biomass = sd(na.omit(total_mass)) / mean(na.omit(total_mass)),
            mean_C = log(mean(na.omit(pct_C))),
            mean_P = log(mean(na.omit(ppm_P))),
            mean_N = log(mean(na.omit(pct_N))),
            mean_K = log(mean(na.omit(ppm_K))),
            mean_pH = log(mean(na.omit(pH))),
            managed = unique(managed),
            burned = unique(burned),
            grazed = unique(grazed),
            anthropogenic = unique(anthropogenic),
            MAP_v2 = unique(MAP_v2),
            MAT_v2 = unique(MAT_v2)) %>%
   filter(site_code %in% full_table$site_code)

prodmod <- lm(log(spatial_mean_biomass) ~ mean_C + mean_P + mean_K + mean_N + mean_pH + MAP_v2 + MAT_v2 + managed + burned + grazed + anthropogenic ,
  data = full_table)
summary(prodmod)

plot_model(prodmod, type = "std")$data %>%
  mutate(alpha_val = as.numeric(if_else(p.value < 0.05, 1, .3))) %>%
  ggplot(aes(x = estimate,
             y = term,
             color = estimate > 0,
             fill = estimate > 0,
             alpha = I(alpha_val))) + 
  geom_vline(xintercept = 0) +
  geom_segment(aes(x = conf.low,
                 xend = conf.high,
                 y = term,
                 yend = term),
               size = 1) +
  geom_point(size = 5, pch = 21, color = "black") +
  theme_bw() +
  xlab("Standardized Coefficient Estimate") +
  ylab("Coefficient Name") +
  guides(color = FALSE,
         fill = FALSE) +
  ggtitle("Estimated Biomass")
```


```{r}

prodmod <- lm(log(spatial_mean_biomass) ~ scale(mean_C) + scale(mean_P) + scale(mean_K) + scale(mean_N) + scale(mean_pH) + scale(MAP_v2) + scale(MAT_v2) + managed + burned + grazed + anthropogenic ,
  data = biomass_model)
plot(prodmod)
summary(prodmod)

testdat <- plot.descriptions %>% 
        filter(year_trt == 0) %>% 
        mutate(
          site_code = site_code,
            log_bmass = log(total_mass),
            mean_C = log(pct_C),
            mean_P = log(ppm_P),
            mean_N = log(pct_N),
            mean_K = log(ppm_K),
            mean_pH = log(pH))

pred_biomass = predict(prodmod, newdata = testdat)

testdat$pred_biomass = exp(pred_biomass)

biomass_test <- testdat %>% 
  select(site_code, pred_biomass) %>%
  group_by(site_code) %>%
  summarise(test_mean_biomass = mean(na.omit(pred_biomass)),
            test_cv_biomass = sd(na.omit(pred_biomass)) / mean(na.omit(pred_biomass)))

biomass_test <- left_join(biomass_test,full_table)
cv_test_mod <- lm(scale(log(spatial_cv_biomass)) ~ scale(log(test_cv_biomass)) + managed + burned + grazed + anthropogenic + scale(MAP_v2) + scale(MAT_v2), data = biomass_test)

summary(cv_test_mod)
plot(cv_test_mod)

plot_model(cv_test_mod)$data %>%
  mutate(alpha_val = as.numeric(if_else(p.value < 0.05, 1, .3))) %>%
  ggplot(aes(x = estimate,
             y = term,
             color = estimate > 0,
             fill = estimate > 0,
             alpha = I(alpha_val))) + 
  geom_vline(xintercept = 0) +
  geom_segment(aes(x = conf.low,
                 xend = conf.high,
                 y = term,
                 yend = term),
               size = 1) +
  geom_point(size = 5, pch = 21, color = "black") +
  theme_bw() +
  xlab("Standardized Coefficient Estimate") +
  ylab("Coefficient Name") +
  guides(color = FALSE,
         fill = FALSE) +
  ggtitle("Estimated Productivity")
```

